---
title: "Inference of Co-Occurrence Networks"
author: "Sarah HUET"
date: "2021 M03 10"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(phyloseq)
library(PLNmodels)
library(tibble)
library(stringr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(gdata)


# load work space data
load(".RData")

# remove all temporary objects from your environment
rm(list = names(.GlobalEnv)[grep("tmp",names(.GlobalEnv))])

```

To identify interacting OTUs and to explore to which extent the altered community properties influenced interactions between OTUs, four co-occurrence networks were constructed across all microcosm samples (i.e., 121 samples) and based on the most abundant OTUs using the same filter as described above (i.e., 258 OTUs). Networks were inferred using a sparse multivariate Poisson log-normal (PLN) model with a latent Gaussian layer and an observed Poisson layer using the PLNmodels package (Chiquet et al., 2019) with an offset corresponding to the number of reads in each sample. Four PLN models were used either without covariate (Model 0) or with each one of the altered community properties (i.e., diversity, composition or density) as a qualitative covariate (Models 1, 2 and 3, respectively), based on the following PLN models (Equations 5 to 8). 
Considering that an OTU of abundance Y, in any l sample follows a Poisson distribution of parameter $\Lambda$ as $Y\sim\mathcal{P}(\mathrm{\Lambda})$, we used the following models:

$$\log(\Lambda_l) = o_l + \mu + Z_l,\ Z_{{l}_{1\le l\le121}} \mathrm{\ iid} \sim \mathcal{N}(0,\sigma^2) \ (5;\ M0)$$

$$\log(\Lambda_{al}) = o_{al} + \mu + \alpha_{a} + Z_{al},\ Z_{{al}_{1\le l\le121}} \mathrm{\ iid}\sim\mathcal{N}(0,\sigma^2) \ (6;M1)$$

$$\log(\Lambda_{bl}) = o_{bl} + \mu + \beta_{b} + Z_{bl},\ Z_{{bl}_{1\le l\le121}} \mathrm{\ iid}\sim\mathcal{N}(0,\sigma^2) \ (7;M2)$$
$$\log(\Lambda_{dl}) = o_{dl} + \mu + \delta_{d} + Z_{dl},\ Z_{{dl}_{1\le l\le121}} \mathrm{\ iid}\sim\mathcal{N}(0,\sigma^2) \ (8;M3)$$

where $l=\left\{1,\ldots,121\right\}$ represents the samples, $a=\left\{1,2\right\}$ represents the diversity, $b=\left\{1,2\right\}$ represents the composition, $d=\left\{1,2,3\right\}$ represents the density, $\alpha$, $\beta$, $\delta$ are the fixed effects of the diversity, composition and density of the manipulated communities, respectively, o is the offset for each sample calculated as the log of the sample read sum and Z is the random sampling effect modelling the data overdispersion.
The best network for each model was selected using a Stability Approach to Regularization Selection (StARS) (Liu et al., 2010). The nodes and links of the four networks were compared using the ggVennDiagram function from the ggVennDiagram package (version 1.2.2). For visualization purposes, only the network inferred from the M0 model was visualized using the Interactive Tree of Life (iTOL) webserver (Letunic & Bork, 2019).


# Networks

## Model without covariate

```{r M0}

tmp_ps = ps_16S_fltr

# calculate offset
tmp_otu_table = t(tmp_ps@otu_table@.Data)
tmp_Offset <- compute_offset(tmp_otu_table, offset = "TSS")
# covariable dataframe
tmp_covariables = data.frame(tmp_ps@sam_data)
tmp_covariables = tmp_covariables[,c("sample","treatment")]
# prepare dataframe
tmp_df <- prepare_data(counts = tmp_otu_table,covariates =  tmp_covariables)
tmp_df$Offset <- tmp_Offset

# calcule network
tmp_network <- PLNnetwork(Abundance ~ 1 + offset(log(Offset)), data = tmp_df)

# obtain best model
tmp_model.StARS <- tmp_network$getBestModel("StARS")

# graph adjacency matrix (value = partial correlation value)
tmp_adjacency.StARS <- as.matrix(tmp_model.StARS$latent_network())
# remove redundant data from the matrix
diag(tmp_adjacency.StARS)=0
tmp_adj=tmp_adjacency.StARS
tmp_adj[lower.tri(tmp_adj,diag=T)]=0
# define a partial correlation threshold
tmp_thrld = 0
# extract edges above a threshold and involved nodes
tmp_edges <- which(abs(tmp_adj)>tmp_thrld,arr.ind=TRUE)
tmp_edges_val = tmp_adj[tmp_edges]
tmp_edges_otu <- tibble("source"=rownames(tmp_adj)[tmp_edges[,1]],
                        "target"=colnames(tmp_adj)[tmp_edges[,2]],
                        "value"=tmp_edges_val)
# name edges
tmp_edges_otu[,"edge_name"] <- paste0(apply(tmp_edges_otu[,c("source","target")],1,sort)[1,],
                                      "__",apply(tmp_edges_otu[,c("source","target")],1,sort)[2,])
# extract involved nodes
tmp_nodes <- unique(c(tmp_edges_otu$source,tmp_edges_otu$target))

# rename model and dataframes
gdata::mv("tmp_model.StARS",paste0("network_M","0"))
gdata::mv("tmp_edges_otu",paste0("network_M","0","_edges"))
gdata::mv("tmp_nodes",paste0("network_M","0","_nodes"))


```


```{r itol_connections}

tmp_itol = tmp_edges_otu

tmp_itol[,"color"] <- ifelse(tmp_itol$value > 0, "#3399FF","#FF3366")

tmp_itol[,"style"] <- rep("normal",nrow(tmp_itol))

tmp_itol$value <- abs(tmp_itol$value)

# extrcat data
write.csv(tmp_itol,"iTOL/tmp_itol_tree_data.csv",row.names = F,quote = F)


```

## Models with covariate

```{r M1_M2_M3}

tmp_ps = ps_16S_fltr

tmp_factor_list = c("incub","broth","density")

for (tmp_i in 1:length(tmp_factor_list)) {
  
  # select a covariate
  tmp_factor = tmp_factor_list[tmp_i]
  
  # calculate offset
  tmp_otu_table = t(tmp_ps@otu_table@.Data)
  tmp_Offset <- compute_offset(tmp_otu_table, offset = "TSS")
  # covariable dataframe
  tmp_covariables = data.frame(tmp_ps@sam_data)
  tmp_covariables = tmp_covariables[,c("sample","treatment",tmp_factor)]
  # prepare dataframe
  tmp_df <- prepare_data(counts = tmp_otu_table,covariates =  tmp_covariables)
  tmp_df$Offset <- tmp_Offset
  # calcule network
  tmp_network <- PLNnetwork(Abundance ~ 0+tmp_df[,tmp_factor] + offset(log(Offset)), data = tmp_df)
  # obtain best model
  tmp_model.StARS <- tmp_network$getBestModel("StARS")
  
  # graph adjacency matrix (value = partial correlation value)
  tmp_adjacency.StARS <- as.matrix(tmp_model.StARS$latent_network())
  # remove redundant data from the matrix
  diag(tmp_adjacency.StARS)=0
  tmp_adj=tmp_adjacency.StARS
  tmp_adj[lower.tri(tmp_adj,diag=T)]=0
  # define a partial correlation threshold
  tmp_thrld = 0
  # extract edges above a threshold and involved nodes
  tmp_edges <- which(abs(tmp_adj)>tmp_thrld,arr.ind=TRUE)
  tmp_edges_val = tmp_adj[tmp_edges]
  tmp_edges_otu <- tibble("source"=rownames(tmp_adj)[tmp_edges[,1]],
                          "target"=colnames(tmp_adj)[tmp_edges[,2]],
                          "value"=tmp_edges_val)
  # name edges
  tmp_edges_otu[,"edge_name"] <- paste0(apply(tmp_edges_otu[,c("source","target")],1,sort)[1,],
                                        "__",apply(tmp_edges_otu[,c("source","target")],1,sort)[2,])
  # extract involved nodes
  tmp_nodes <- unique(c(tmp_edges_otu$source,tmp_edges_otu$target))

  # rename model and dataframes
  gdata::mv("tmp_model.StARS",paste0("network_M",tmp_i))
  gdata::mv("tmp_edges_otu",paste0("network_M",tmp_i,"_edges"))
  gdata::mv("tmp_nodes",paste0("network_M",tmp_i,"_nodes"))

}

```

# Venn diagrams


```{r Venn_Diagram_edges}

library(ggVennDiagram)

tmp_venn <- list("all"=network_M0_edges$edge_name,
                 "div"=network_M1_edges$edge_name,
                 "compo"=network_M2_edges$edge_name,
                 "dens"=network_M3_edges$edge_name)

# 4D Venn diagram
venn_edges <- ggVennDiagram(tmp_venn, color = "black", lwd = 0.8, lty = 1)
venn_edges

```

```{r Venn_Diagram_nodes}

library(ggVennDiagram)

tmp_venn <- list("all"=network_M0_nodes,
                 "div"=network_M1_nodes,
                 "compo"=network_M2_nodes,
                 "dens"=network_M3_nodes)

# 4D Venn diagram
venn_nodes <- ggVennDiagram(tmp_venn, color = "black", lwd = 0.8, lty = 1)
venn_nodes

```

```{r barplot}

# extract Venn intersection
tmp = venn_edges[["plot_env"]][["data"]]@region

# number of positive and negative per Venn intersection
for (tmp_i in tmp$name) {
  
  tmp_pos = length(which(unlist(tmp$item[tmp$name == tmp_i]) %in%
                       network_M0_edges$edge_name[network_M0_edges$value >0]))
  
  tmp_neg = length(which(unlist(tmp$item[tmp$name == tmp_i]) %in%
                       network_M0_edges$edge_name[network_M0_edges$value <0]))
  
  tmp[tmp$name == tmp_i,"pos"] <- tmp_pos
  tmp[tmp$name == tmp_i,"neg"] <- tmp_neg
}

# percent
tmp <- tmp %>%
  mutate(
    pos_pc = pos / length(which(network_M0_edges$value >0)) *100,
    neg_pc = -(neg / length(which(network_M0_edges$value <0)) *100)
  ) %>%
  pivot_longer(.,cols=c(pos_pc,neg_pc),names_to = "link_value",)

# barplot
tmp %>%
 filter(grepl("all",name)) %>%
 ggplot() +
  aes(x = name, y = value, fill = link_value) +
  geom_col() +
  coord_flip() +
  theme_light()



```





