---
title: "1.2.Upset"
author: "Sarah HUET"
date: "2021 M12 22"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# remove all temporary objects from your environment
rm(list = names(.GlobalEnv)[grep("tmp",names(.GlobalEnv))])

# load libraries
library(phyloseq)
library(tibble)
library(tidyr)

```

# Comparison : subset sample type before/after filter most abundant OTUs

## subset before filter

```{r subset}

tmp_ps0 = ps_16S

tmp_ps_micro <- prune_samples(tmp_ps0@sam_data$sample_type == "microcosm",tmp_ps0)
tmp_ps_micro <- prune_taxa(taxa_sums(tmp_ps_micro) > 0, tmp_ps_micro)

tmp_ps_susp <- prune_samples(tmp_ps0@sam_data$sample_type == "suspension",tmp_ps0)
tmp_ps_susp <- prune_taxa(taxa_sums(tmp_ps_susp) > 0, tmp_ps_susp)


```

```{r filter}

for (tmp_sample_type in c("microcosm","suspension")) {
  
  if (tmp_sample_type == "microcosm") {
    tmp_ps = tmp_ps_micro
  } else {
    tmp_ps = tmp_ps_susp
  }
  
  ##### frequency -----

  # calculate OTU frequency
  tmp_df_otu <- as.data.frame(otu_table(tmp_ps))
  tmp_df_otu_freq <- apply(tmp_df_otu, 2, FUN=function(x) x/sum(x)*100)
  # apply a minimum frequency threshold
  tmp_freq_thld = 0.25
  tmp <- apply(tmp_df_otu_freq, 1, FUN=function(x) sum(x>=(tmp_freq_thld)))
  # select OTUs above frequency threshold
  tmp_otus_F1 <- rownames(tmp_df_otu[-which(tmp==0),])
  # subset selected OTUs
  tmp_ps_filter1 <- prune_taxa(taxa_names(tmp_ps) %in% tmp_otus_F1, tmp_ps)
  
  
  
  ##### prevalence -----
  
  tmp_ps = tmp_ps_filter1 
  # calculate OTUs prevalence in treatment (ttt)
  tmp_df <- psmelt(tmp_ps)
  tmp_otu_prev_ttt <- data.frame(matrix(ncol=length(unique(tmp_df$treatment)),
                                       nrow=length(unique(tmp_df$OTU)), 
                                       dimnames=list(unique(tmp_df$OTU),
                                                     unique(tmp_df$treatment))))
  for (i in unique(tmp_df$OTU)) {
    for (j in unique(tmp_df$treatment)) {
      tmp_otu_prev_ttt[i,j] <- sum(tmp_df$Abundance[tmp_df$OTU == i & tmp_df$treatment == j] > 0,
                                   na.rm = T) / length(tmp_df$Sample[tmp_df$OTU == i & tmp_df$treatment == j]) *100
    }
    
  } 
  rm(i,j)
  # calculate maximum OTUs prevalence in treatment
  tmp <- apply(tmp_otu_prev_ttt,1, FUN=function(x) max(x))
  # select OTUs above a minimum prevalence in treatment threshold set to 60% 
  tmp_otus_F2 <- rownames(tmp_otu_prev_ttt[which(tmp >= 60),])
  # subset selected OTUs
  tmp_ps_filter2 <- prune_taxa(taxa_names(tmp_ps) %in% tmp_otus_F2, tmp_ps)
  
  if (tmp_sample_type == "microcosm") {
    tmp_ps_micro_fltr = tmp_ps_filter2
  } else {
    tmp_ps_susp_fltr = tmp_ps_filter2
  }

}

rm(tmp_df,tmp_df_otu,tmp_df_otu_freq,tmp_otu_prev_ttt,tmp,tmp_freq_thld,tmp_otus_F1,tmp_otus_F2,tmp_sample_type)
rm(tmp_ps_filter1,tmp_ps_filter2)
rm(tmp_ps,tmp_ps0)



```

## subset after filter

```{r filter}

tmp_ps0 = tmp_ps = ps_16S

##### frequency -----

# calculate OTU frequency
tmp_df_otu <- as.data.frame(otu_table(tmp_ps))
tmp_df_otu_freq <- apply(tmp_df_otu, 2, FUN=function(x) x/sum(x)*100)
# apply a minimum frequency threshold
tmp_freq_thld = 0.25
tmp <- apply(tmp_df_otu_freq, 1, FUN=function(x) sum(x>=(tmp_freq_thld)))
# select OTUs above frequency threshold
tmp_otus_F1 <- rownames(tmp_df_otu[-which(tmp==0),])
# subset selected OTUs
tmp_ps_filter1 <- prune_taxa(taxa_names(tmp_ps) %in% tmp_otus_F1, tmp_ps)

##### prevalence -----

tmp_ps = tmp_ps_filter1 
# calculate OTUs prevalence in treatment (ttt)
tmp_df <- psmelt(tmp_ps)
tmp_otu_prev_ttt <- data.frame(matrix(ncol=length(unique(tmp_df$treatment)),
                                     nrow=length(unique(tmp_df$OTU)), 
                                     dimnames=list(unique(tmp_df$OTU),
                                                   unique(tmp_df$treatment))))
for (i in unique(tmp_df$OTU)) {
  for (j in unique(tmp_df$treatment)) {
    tmp_otu_prev_ttt[i,j] <- sum(tmp_df$Abundance[tmp_df$OTU == i & tmp_df$treatment == j] > 0,
                                 na.rm = T) / length(tmp_df$Sample[tmp_df$OTU == i & tmp_df$treatment == j]) *100
  }
  
} 
rm(i,j)
# calculate maximum OTUs prevalence in treatment
tmp <- apply(tmp_otu_prev_ttt,1, FUN=function(x) max(x))
# select OTUs above a minimum prevalence in treatment threshold set to 60% 
tmp_otus_F2 <- rownames(tmp_otu_prev_ttt[which(tmp >= 60),])
# subset selected OTUs
tmp_ps_filter2 <- prune_taxa(taxa_names(tmp_ps) %in% tmp_otus_F2, tmp_ps)

tmp_ps_fltr_all = tmp_ps_filter2

rm(tmp_df,tmp_df_otu,tmp_df_otu_freq,tmp_otu_prev_ttt,tmp,tmp_freq_thld,tmp_otus_F1,tmp_otus_F2)
rm(tmp_ps_filter1,tmp_ps_filter2)
rm(tmp_ps,tmp_ps0)

```

```{r subset}

tmp_ps0 = tmp_ps_fltr_all

tmp_ps_fltr_micro <- prune_samples(tmp_ps0@sam_data$sample_type == "microcosm",tmp_ps0)
tmp_ps_fltr_micro <- prune_taxa(taxa_sums(tmp_ps_fltr_micro) > 0, tmp_ps_fltr_micro)

tmp_ps_fltr_susp <- prune_samples(tmp_ps0@sam_data$sample_type == "suspension",tmp_ps0)
tmp_ps_fltr_susp <- prune_taxa(taxa_sums(tmp_ps_fltr_susp) > 0, tmp_ps_fltr_susp)

rm(tmp_ps0)

```

## upset

```{r upset data}

tmp_OTU_all = unique(c(taxa_names(tmp_ps_fltr_all),
                       taxa_names(tmp_ps_fltr_micro),
                       taxa_names(tmp_ps_fltr_susp),
                       taxa_names(tmp_ps_micro_fltr),
                       taxa_names(tmp_ps_susp_fltr)))


tmp_upset <- tibble("OTU"=tmp_OTU_all)

tmp_upset[tmp_upset$OTU %in% taxa_names(tmp_ps_fltr_all),"fltr_all"] <- 1
tmp_upset[tmp_upset$OTU %in% taxa_names(tmp_ps_fltr_micro),"fltr_micro"] <- 1
tmp_upset[tmp_upset$OTU %in% taxa_names(tmp_ps_fltr_susp),"fltr_susp"] <- 1
tmp_upset[tmp_upset$OTU %in% taxa_names(tmp_ps_micro_fltr),"micro_fltr"] <- 1
tmp_upset[tmp_upset$OTU %in% taxa_names(tmp_ps_susp_fltr),"susp_fltr"] <- 1

tmp_upset[is.na(tmp_upset)] <- 0

```

```{r upset metadata}

tmp_ps = prune_taxa(taxa_names(ps_16S) %in% tmp_OTU_all, ps_16S)


# load and subset otu table
library(readr)
tmp_otu_table <- read_delim("Data/otu_table.csv",
                            delim = ",", escape_double = FALSE, trim_ws = TRUE)
tmp_otu_table = tmp_otu_table[tmp_otu_table$OTU %in% tmp_OTU_all,]


# calculate prevalence and abundance
tmp_prev = apply(tmp_ps@otu_table,
                 MARGIN = ifelse(taxa_are_rows(tmp_ps),yes = 1, no = 2),
                 FUN=function(x) sum(x>0)/nsamples(tmp_ps)*100)

tmp_upset_data = tibble("OTU"=taxa_names(tmp_ps),
                        "ASV total abundance"=taxa_sums(tmp_ps),
                        "ASV total abundance (log10)"=log10(taxa_sums(tmp_ps)),
                        "Prevalence"=tmp_prev)
tmp_upset_data = left_join(tmp_upset_data,tmp_upset,by="OTU")

tmp_upset_metadata = left_join(tmp_otu_table,tmp_upset_data,by="OTU")

# extract metadata
write.csv(tmp_upset_metadata,file = "Upset/upset_comparison_filters_sample_type_metadata.csv",row.names = F)

# check OTU nowhere and everywhere
tmp_otu_nowhere = tmp_upset[which(rowSums(tmp_upset[,-1]) <= 0),-1]
tmp_otu_everywhere = tmp_upset[which(rowSums(tmp_upset[,-1]) == ncol(tmp_upset[,-1])),-1]

```

# comparison : original soil and five suspensions

```{r wrangle data}

tmp_ps0 = ps_16S

# subset treatment
tmp_ps = prune_samples(tmp_ps0@sam_data$sample_type %in% c("original_soil","suspension"),tmp_ps0)
# merged by treatment (induce NA in sample data)
tmp_ps_merged = merge_samples(tmp_ps,group = tmp_ps@sam_data$treatment)
# extract ps data
tmp_psmelt = psmelt(tmp_ps_merged)

```

```{r upset}

tmp_upset = tibble("OTU"=tmp_psmelt$OTU,
                   "treatment"=tmp_psmelt$Sample,
                   "abundance"=tmp_psmelt$Abundance)
# transform count
tmp_upset$abundance[tmp_upset$abundance >0] <- 1

# pivot table
tmp_upset = tmp_upset %>% pivot_wider(names_from = treatment, values_from = abundance)

# check OTU
tmp_OTU_all = tmp_upset$OTU
tmp_otu_nowhere = tmp_upset[which(rowSums(tmp_upset[,-1]) <= 0),-1]
tmp_otu_everywhere = tmp_upset[which(rowSums(tmp_upset[,-1]) == ncol(tmp_upset[,-1])),-1]

```

```{r upset metadata}

tmp_ps = prune_taxa(taxa_names(ps_16S) %in% tmp_OTU_all, ps_16S)


# load and subset otu table
library(readr)
tmp_otu_table <- read_delim("Data/otu_table.csv",
                            delim = ",", escape_double = FALSE, trim_ws = TRUE)
tmp_otu_table = tmp_otu_table[tmp_otu_table$OTU %in% tmp_OTU_all,]

# calculate prevalence and abundance
tmp_prev = apply(tmp_ps@otu_table,
                 MARGIN = ifelse(taxa_are_rows(tmp_ps),yes = 1, no = 2),
                 FUN=function(x) sum(x>0)/nsamples(tmp_ps)*100)

tmp_upset_data = tibble("OTU"=taxa_names(tmp_ps),
                        "ASV total abundance"=taxa_sums(tmp_ps),
                        "ASV total abundance (log10)"=log10(taxa_sums(tmp_ps)),
                        "Prevalence"=tmp_prev)
tmp_upset_data = left_join(tmp_upset_data,tmp_upset,by="OTU")

# join tibbles
tmp_upset_metadata = left_join(tmp_otu_table,tmp_upset_data,by="OTU")

# extract metadata
write.csv(tmp_upset_metadata,
          file = "Upset/upset_comparison_original_soil_suspensions_metadata.csv",
          row.names = F)

```

# comparison : treatments

```{r wrangle data}

tmp_ps0 = tmp_ps = ps_16S

# merged by treatment (induce NA in sample data)
tmp_ps_merged = merge_samples(tmp_ps,group = tmp_ps@sam_data$treatment)
# extract ps data
tmp_psmelt = psmelt(tmp_ps_merged)

```

```{r upset}

tmp_upset = tibble("OTU"=tmp_psmelt$OTU,
                   "treatment"=tmp_psmelt$Sample,
                   "abundance"=tmp_psmelt$Abundance)
# transform count
tmp_upset$abundance[tmp_upset$abundance >0] <- 1

# pivot table
tmp_upset = tmp_upset %>% pivot_wider(names_from = treatment, values_from = abundance)

# check OTU
tmp_OTU_all = tmp_upset$OTU
tmp_otu_nowhere = tmp_upset[which(rowSums(tmp_upset[,-1]) <= 0),-1]
tmp_otu_everywhere = tmp_upset[which(rowSums(tmp_upset[,-1]) == ncol(tmp_upset[,-1])),-1]

```

```{r upset metadata}

tmp_ps = ps_16S


# load and subset otu table
library(readr)
tmp_otu_table <- read_delim("Data/otu_table.csv",
                            delim = ",", escape_double = FALSE, trim_ws = TRUE)
tmp_otu_table = tmp_otu_table[tmp_otu_table$OTU %in% tmp_OTU_all,]

# calculate prevalence and abundance
tmp_prev = apply(tmp_ps@otu_table,
                 MARGIN = ifelse(taxa_are_rows(tmp_ps),yes = 1, no = 2),
                 FUN=function(x) sum(x>0)/nsamples(tmp_ps)*100)

tmp_upset_data = tibble("OTU"=taxa_names(tmp_ps),
                        "ASV total abundance"=taxa_sums(tmp_ps),
                        "ASV total abundance (log10)"=log10(taxa_sums(tmp_ps)),
                        "Prevalence"=tmp_prev)
tmp_upset_data = left_join(tmp_upset_data,tmp_upset,by="OTU")

# join tibbles
tmp_upset_metadata = left_join(tmp_otu_table,tmp_upset_data,by="OTU")

# extract metadata
write.csv(tmp_upset_metadata,file = "Upset/upset_comparison_treatments_metadata.csv",row.names = F)

```

## count

We want the % of OTU in suspensions detected in coalesced communities
among the OTU detected in each susp, how many in detected in the corresponding coal ?

```{r}

tmp_data = tmp_upset %>%
  pivot_longer(!OTU,names_to = "treatment",values_to = "detection")

tmp_manips = c("MAC_a1","MAC_a2","PEB_a1","PEB_a2")

for (tmp_manip in tmp_manips) {#tmp_manip = tmp_manips[1]
  ## OTU in susp
  tmp_OTU_susp_C = tmp_data$OTU[tmp_data$treatment == "susp_C" & 
                                  tmp_data$detection == 1]
  
  tmp_OTU_susp_manip = tmp_data$OTU[tmp_data$treatment == paste0("susp_",tmp_manip) & 
                                  tmp_data$detection == 1]
  ## OTUs in coal
  tmp_OTU_coal = unique(tmp_data$OTU[tmp_data$treatment %in% paste0(tmp_manip,"d",rep(1:3,1)) & 
                                  tmp_data$detection == 1])
  
  ## percent coal / susp_C
  tmp = intersect(tmp_OTU_susp_C,tmp_OTU_coal)
  tmp_percent = round(length(tmp) / length(tmp_OTU_susp_C) *100,1)
  print(paste0("Among the OTUs detected in the control suspension, ",tmp_percent,
               " % were also detected in the coalescence treatment with the ",tmp_manip," suspension"))
  tmp = intersect(tmp_OTU_susp_manip,tmp_OTU_coal)
  tmp_percent = round(length(tmp) / length(tmp_OTU_susp_manip) *100,1)
  print(paste0("Among the OTUs detected in the manipulated suspension ",tmp_manip,", ",tmp_percent,
               " % were also detected in the corresponding coalescence treatment"))
  
}


```





# plot

faire un plot où x = OTU présentes dans les suspensions MAC et PEB et y = leur abondance relative dans le sol original

```{r data}

tmp_ps0 = ps_16S

# which OTU are selected in MAC an PEB suspensions
tmp_ps = tmp_ps0
tmp_ps = prune_samples(tmp_ps@sam_data$sample_type == "suspension", tmp_ps)
tmp_ps = prune_samples(tmp_ps@sam_data$treatment != "susp_C",tmp_ps)
tmp_ps = prune_taxa(taxa_sums(tmp_ps) > 0, tmp_ps)

tmp_selected_otu = taxa_names(tmp_ps)

# what abundance those OTU have in T0
tmp_ps = tmp_ps0
tmp_ps = prune_samples(tmp_ps@sam_data$treatment == "T0",tmp_ps)
tmp_ps <- transform_sample_counts(tmp_ps,function(x) x / sum(x))
tmp_ps = prune_taxa(taxa_names(tmp_ps) %in% tmp_selected_otu, tmp_ps)

tmp_selected_otu = tibble("OTU"=tmp_selected_otu,
                          "rel_ab"=taxa_sums(tmp_ps))

```

```{r plot}

ggplot(tmp_selected_otu,aes(x=reorder(OTU,desc(rel_ab)),
                            y=rel_ab))+
  geom_col()+
  theme_bw() +
  labs(x = "OTU present in MAC and PEB suspensions",
       y = "OTU average relative abundance in original soil samples")+
  theme(axis.text.x  = element_text(vjust=0.5, size=5,angle = 90))


```

```{r info about T0 samples}

tmp_ps0 = ps_16S
tmp_ps = prune_samples(tmp_ps0@sam_data$treatment == "T0",tmp_ps0)
tmp_ps = prune_taxa(taxa_sums(tmp_ps) > 0, tmp_ps)

sum(sample_sums(tmp_ps))
ntaxa(tmp_ps)
max(taxa_sums(tmp_ps))
min(taxa_sums(tmp_ps))
mean(taxa_sums(tmp_ps))
sd(taxa_sums(tmp_ps))
median(taxa_sums(tmp_ps))

tmp_ps <- transform_sample_counts(tmp_ps,function(x) x / sum(x))

sum(sample_sums(tmp_ps))
ntaxa(tmp_ps)
max(taxa_sums(tmp_ps))
min(taxa_sums(tmp_ps))
mean(taxa_sums(tmp_ps))
sd(taxa_sums(tmp_ps))
median(taxa_sums(tmp_ps))

tmp_selected_otu = tibble("OTU"=taxa_names(tmp_ps),
                          "rel_ab"=taxa_sums(tmp_ps))

tmp_selected_otu <- tmp_selected_otu[tmp_selected_otu$rel_ab >= 0.01,]

# plot

p <- ggplot(tmp_selected_otu,aes(x=reorder(OTU,desc(rel_ab)),
                                 y=rel_ab))+
  geom_col()+
  theme_bw() +
  labs(x = "OTU present in Original soil samples",
       y = "OTU average relative abundance in original soil samples")+
  theme(axis.text.x  = element_text(vjust=0.5, size=5,angle = 90))

p

```





















