---
title: "1_Load_data"
author: "Sarah HUET"
date: "2021 M03 10"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script take biom file (QIIME pipeline output) and mapping file to gather data in phyloseq objects:
     - ps_16S_raw & ps_18S_raw: phyloseq objects with raw data, for 16S rDNA and 18S rDNA sequences respectively
     - ps_16S & ps_18S: phyloseq objects with cleaned data, for 16S rDNA and 18S rDNA sequences respectively
     - ps_16S_fltr & ps_18S_most_abund: phyloseq object with filtered OTUs, for 16S rDNA and 18S rDNA sequences respectively

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(phyloseq)
library(readr)
library(readxl)
library(dplyr)
library(phytools)

# remove all temporary objects from your environment
rm(list = names(.GlobalEnv)[grep("tmp",names(.GlobalEnv))])

```

# Load raw biom file and clean data

```{r load additional data}

Colors_treatments <- read_excel("Data/Colors_treatments.xlsx")

```


```{r load raw biom file}

# import biom file
tmp_ps = import_biom("Data/otu_table.biom")
# remove extra taxa rank
tmp_ps@tax_table <- tmp_ps@tax_table[,1:7]
# set taxa rank names
colnames(tmp_ps@tax_table) <- c("Kingdom","Phylum","Class","Order","Family","Genus","Species")
# import mapping file
metadata_sample <- read_excel("Data/Metadata_xp2_coal_sample.xlsx")
tmp_design = sample_data(metadata_sample)
sample_names(tmp_design) <- tmp_design$sample
# import tree
tmp_tree = read.newick("Data/phylogeny.tre")
#merge onto one phyloseq object
ps_16S_raw = merge_phyloseq(tmp_ps,tmp_design,tmp_tree)

# final ps_16S_raw: 155 samples & 3227 OTUs

rm(list = names(.GlobalEnv)[grep("tmp_",names(.GlobalEnv))])

```

```{r clean dataset}

# remove samples with seq count < tmp_rare_thld
tmp_rare_thld = 8000
ps_16S <- prune_samples(sample_sums(ps_16S_raw) > tmp_rare_thld, ps_16S_raw )
ps_16S <- prune_taxa(taxa_sums(ps_16S) > 0, ps_16S)

# final ps_16S 8000 : 146 samples & 3227 OTUs

#remove chloroplasts /!\ only with greengenes
#tmp_chloroplast = taxa_names(ps_16S_raw)[which(ps_16S_raw@tax_table[,"Class"] == "c__Chloroplast")]
#ps_16S <- prune_taxa(!(taxa_names(ps_16S) %in% tmp_chloroplast), ps_16S)

# check which OTUs where removed
setdiff(taxa_names(ps_16S_raw),taxa_names(ps_16S))
# check which samples where removed
setdiff(sample_names(ps_16S_raw),sample_names(ps_16S))

```

## Split microcosm samples from suspension samples


```{r microcosms}

ps_16S_micro <- prune_samples(ps_16S@sam_data$sample_type == "microcosm",ps_16S)
ps_16S_micro <- prune_taxa(taxa_sums(ps_16S_micro) > 0, ps_16S_micro)

# final ps_16S_micro: 399 samples & 5500 OTUs

```

```{r original soil }

ps_16S_T0 <- prune_samples(ps_16S@sam_data$sample_type == "original_soil",ps_16S)
ps_16S_T0 <- prune_taxa(taxa_sums(ps_16S_T0) > 0, ps_16S_T0)

# final ps_16S_susp: 399 samples & 5500 OTUs

```

```{r suspensions }

ps_16S_susp <- prune_samples(ps_16S@sam_data$sample_type == "suspension",ps_16S)
ps_16S_susp <- prune_taxa(taxa_sums(ps_16S_susp) > 0, ps_16S_susp)

# final ps_16S_susp: 399 samples & 5500 OTUs

# only manipulated suspensions
ps_16S_susp_manip <- prune_samples(ps_16S_susp@sam_data$treatment != "susp_C",ps_16S_susp)
ps_16S_susp_manip <- prune_taxa(taxa_sums(ps_16S_susp_manip) > 0, ps_16S_susp_manip)


```


## Filter most abundant OTUs

```{r freq filter}

tmp_ps = ps_16S_micro
# calculate OTU frequency
tmp_df_otu <- as.data.frame(otu_table(tmp_ps))
tmp_df_otu_freq <- apply(tmp_df_otu, 2, FUN=function(x) x/sum(x)*100)
# apply a minimum frequency threshold
tmp_freq_thld = 0.1
tmp <- apply(tmp_df_otu_freq, 1, FUN=function(x) sum(x>=(tmp_freq_thld)))
# select OTUs above frequency threshold
#tmp_otus_F1 <- rownames(tmp_df_otu[-which(tmp==0),])
tmp_nb_sample = 10
tmp_otus_F1 <- rownames(tmp_df_otu[which(tmp>=tmp_nb_sample),])
# subset selected OTUs
tmp_ps_fltr1 <- prune_taxa(taxa_names(tmp_ps) %in% tmp_otus_F1, tmp_ps)

```


```{r prevalence filter}
# /!\ takes from 30min to 3h /!\

tmp_ps = tmp_ps_fltr1
# calculate OTUs prevalence in treatment (ttt)
tmp_df <- psmelt(tmp_ps)
tmp_otu_prev_ttt <- data.frame(matrix(ncol=length(unique(tmp_df$treatment)),
                                     nrow=length(unique(tmp_df$OTU)), 
                                     dimnames=list(unique(tmp_df$OTU),
                                                   unique(tmp_df$treatment))))
for (i in unique(tmp_df$OTU)) {
  for (j in unique(tmp_df$treatment)) {
    tmp_otu_prev_ttt[i,j] <- sum(tmp_df$Abundance[tmp_df$OTU == i & tmp_df$treatment == j] > 0,
                                 na.rm = T) / length(tmp_df$Sample[tmp_df$OTU == i & tmp_df$treatment == j]) *100
  }
  
} 
rm(i,j)
# calculate maximum OTUs prevalence in treatment
tmp <- apply(tmp_otu_prev_ttt,1, FUN=function(x) max(x))
# select OTUs above a minimum prevalence in treatment threshold set to 60% 
tmp_otus_F2 <- rownames(tmp_otu_prev_ttt[which(tmp >= 60),])
# subset selected OTUs
tmp_ps_fltr2 <- prune_taxa(taxa_names(tmp_ps) %in% tmp_otus_F2, tmp_ps)

ps_16S_fltr = tmp_ps_fltr2

```


# Measurement of community functions 

```{r import data}

data_function <- read_excel("Data/Metadata_xp2_coal_pool.xlsx")
# remove unused treatments (treatments not in this experimental design)
data_function = data_function[!data_function$pool %in% c(141:155),]

```
