---
title: "2_alpha_div"
author: "Sarah HUET"
date: "2021 M03 10"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# remove all temporary objects from your environment
rm(list = names(.GlobalEnv)[grep("tmp",names(.GlobalEnv))])

# load libraries
library(tidyr)
library(ggplot2)
library(dplyr)
library(agricolae)
library(tibble)

```

# load alpha diversity measurments

Alpha diversity indexes were calculated with an home made Qiime pipeline implemented in Python. Here, we load the resulting dataframe.

```{r wrangle data 16S}

tmp_path_16S = "C:/Users/srhuet/Documents/copie_DDE_20220307/Thesis/EMFEED/BISCAL_xp2_coal/Dalton/out/collate_alpha.d/"

# create a df that gather all alpha div measures
alphadiv_16S_all <- vector("list", 1+length(list.files(path = tmp_path_16S)))

for (i in list.files(path = tmp_path_16S)) {
  
  tmp_txt_path = paste0(tmp_path_16S,i)
  
  tmp_alphadiv <- read.table(tmp_txt_path,sep="\t", header=TRUE)
  tmp_alphadiv <- tmp_alphadiv[tmp_alphadiv$sequences.per.sample == 8000,4:157]
  
  tmp_alphadiv <- as_tibble(tmp_alphadiv) %>% gather("sample", "value")
  tmp_alphadiv[tmp_alphadiv$value == 'n/a',"value"] <- NA
  tmp_alphadiv[,i] <- as.numeric(tmp_alphadiv$value)
  
  alphadiv_16S_all[[1]]$sample <- gsub("X","",tmp_alphadiv$sample)
  alphadiv_16S_all[[length(which(alphadiv_16S_all != "NULL"))+1]] <- tmp_alphadiv[,i]
  
  #next step introduce NA. why ???
  print(i)
  rm(tmp_alphadiv)
}

alphadiv_16S_all <- do.call(cbind,alphadiv_16S_all)
alphadiv_16S_all <- alphadiv_16S_all[complete.cases(alphadiv_16S_all),]
colnames(alphadiv_16S_all) <- gsub(".txt","",colnames(alphadiv_16S_all))
alphadiv_16S_all$sample <- as.numeric(alphadiv_16S_all$sample)

# add the corresponding treatment
alphadiv_16S_all <- full_join(alphadiv_16S_all,metadata_sample,by="sample")

# add color and order
alphadiv_16S_all <- left_join(alphadiv_16S_all,Colors_treatments,by = "treatment")

alphadiv_16S_all$sample <- as.numeric(alphadiv_16S_all$sample)

write.csv(alphadiv_16S_all,"Data/alphadiv_all.csv",row.names = F)

```

# stats

## data
```{r choose your data}

# select which data you want to test

tmp_sample_type = "suspension" #  microcosm| suspension  | modif_susp | all
tmp_index = "simpson_reciprocal" # observed_species / simpson_reciprocal / PD_whole_tree

```

```{r wrangle data}

tmp_alphadiv = alphadiv_16S_all
tmp_alphadiv <- tmp_alphadiv[complete.cases(tmp_alphadiv),]

if (tmp_sample_type == "microcosm") {
  tmp_data <- tmp_alphadiv[tmp_alphadiv$sample_type == tmp_sample_type,] # select microcosm treatments
}
if (tmp_sample_type == "suspension"){
  tmp_data <- tmp_alphadiv[tmp_alphadiv$sample_type %in% c("original_soil","suspension"),]
}
if (tmp_sample_type == "modif_susp"){
  tmp_data <- tmp_alphadiv[tmp_alphadiv$sample_type =="suspension",]
  tmp_data <- tmp_data[tmp_data$treatment != "susp_C",]
}
if (tmp_sample_type == "all"){
  tmp_data <- tmp_alphadiv
}

# select the chosen index
tmp_data <- tmp_data[,c("sample",tmp_index)]
colnames(tmp_data)[2] <- "value"

# concatenate iterations for each sample by mean
tmp_data <- tmp_data %>% 
  dplyr::group_by(sample) %>% 
  dplyr::summarise(value = mean(value))

# add treatment
tmp_data <- left_join(tmp_data,metadata_sample[,c("sample","treatment")])
# add color and order
tmp_data <- left_join(tmp_data,
                      Color_sets[Color_sets$set == "treatments",c("treatment","color","order")],
                      by = "treatment")

```

```{r summarise adiv}

tmp <- tmp_data %>%
      dplyr::group_by(treatment) %>%
  dplyr::summarise(avg = mean(value))

tmp[,"diff_C"] <- 100-(tmp$avg / tmp$avg[tmp$treatment =="susp_C"] *100)

tmp


```

## tests
```{r statistic test}

# linear model
tmp_lm = lm(tmp_data$value ~ tmp_data$treatment, na.action = na.omit )
summary(tmp_lm)
# anova
anova(tmp_lm)
tmp_aov = aov(tmp_lm)
summary(tmp_aov)
# post hoc Tukey test
tmp_comp <- HSD.test(tmp_aov,'tmp_data$treatment',alpha = 0.05,group = T)
# tibble with statistical groups
tmp_stat = tibble("treatment"=rownames(tmp_comp[["groups"]]),
                  "mean"=tmp_comp[["groups"]][["tmp_data$value"]],
                  "stat_groups"=as.character(tmp_comp[["groups"]][["groups"]]))

tmp_stat <- inner_join(tmp_stat,Colors_treatments[,c("treatment","color","order")],by = "treatment")


```

```{r barplot stat with letters, width = 5.5,height = 6}
library(ggplot2)

tmp_ylim = c(0,max(tmp_data$value)*1.2)
#tmp_ylim = c(ceiling(min(tmp_data$value)*0.75),max(tmp_data$value)*1.2)
tmp_ylim_stat = max(tmp_data$value)*1.1

tmp_plot_data = tmp_data
tmp_plot_stat = tmp_stat

tmp_plot_stat[,"stat_y"] <- tmp_ylim_stat

# plot
p <- ggplot(tmp_plot_data, aes(x = reorder(treatment,desc(order)),
                               #y = log10(value))) +
                               y = value)) +
  
  stat_summary(geom = "bar", fun = mean, colour = "#333333",
               linetype = 1,
               fill = levels(reorder(tmp_plot_data$color,desc(tmp_plot_data$order))))+ 
  labs(x="Treatment", y= paste0("\n",tmp_index), title = tmp_index, color = "#333333")+
  coord_flip(ylim = tmp_ylim)+
  #scale_y_log10()+
  theme_bw()+
  theme(title = element_text(face="bold", size=12),
        legend.position = "right",
        axis.title.x = element_text(face="bold", size=10), 
        axis.text.x  = element_text(vjust=0.5, size=8,angle = 0),
        axis.title.y = element_text(face="bold", size=10), 
        axis.text.y  = element_text(vjust=0.5, size=8),
        strip.text = element_text(face="bold", size=8),
        strip.background = element_rect(fill="white", colour="#333333",size=1.5))+
  stat_summary(geom = "linerange", fun.data = mean_se, width=0.5,color = "#333333" )+
  geom_text(data = tmp_plot_stat,
            #mapping = aes(label = stat_groups, y = stat_y*5, x=treatment),
            mapping = aes(label = stat_groups, y = stat_y, x=treatment),
            colour="#333333", size=8,inherit.aes = F, angle = 0,
            position = position_nudge(x = 0))
  
p

fig_div[[paste0(tmp_sample_type,"_",tmp_index)]] <- p
tmp_div[[tmp_index]] <- tmp_data %>% dplyr::group_by(treatment) %>% dplyr::summarise(avg = mean(value))

```




## basic estimators

```{r microcosms}

# mean
tmp_mean = tmp_data %>% group_by(treatment) %>% summarise(mean = mean(value))

## mean C
tmp_mean_C = tmp_mean$mean[tmp_mean$treatment == "C"]

# diff C and min microcosm
100-(min(tmp_mean$mean)/tmp_mean_C*100)

# diff C and max microcosm
max(tmp_mean$mean)/tmp_mean_C*100

# diff and mean microcosms
100-(mean(tmp_mean$mean)/tmp_mean_C*100)

```

```{r susp}

# mean
tmp_mean = tmp_data %>% group_by(treatment) %>% summarise(mean = mean(value))

## mean C
tmp_mean_C = tmp_mean$mean[tmp_mean$treatment == "susp_C"]

## mean susp
tmp_mean_susp = tmp_mean$mean[!tmp_mean$treatment %in% c("T0","susp_C")]
min(tmp_mean_susp)


# diff C and mean susp
100-(mean(tmp_mean_susp)/tmp_mean_C*100)

# diff C and max susp
max(tmp_mean$mean)/tmp_mean_C*100



```



```{r export plot}


ggsave(plot = p,
       filename = "tmp_fig.svg",
       width = 95,height = 110,units = "mm")





library(officer)
library(rvg)

tmp_editable_graph <- dml(ggobj = p)
doc <- read_pptx()
doc <- add_slide(doc)
doc <- ph_with(x = doc, tmp_editable_graph,
               location = ph_location_type(type = "body") )
print(doc, target = "tmp_fig.pptx")

```

from https://ardata-fr.github.io/officeverse/officer-for-powerpoint.html

# SUSPENSIONS t- test vs T0
as there is an unbalanced design and a low number of replicate for suspensions  we'll use welch t-test 

we asked two questions
(1) are the suspensions different from the original soil?
(2) are the manipulated suspensions different between them?

```{r choose your data}

tmp_index = "observed_species" # , PD_whole_tree, simpson_reciprocal

# select which data you want to test
tmp_alphadiv = alphadiv_16S_all
tmp_alphadiv <- tmp_alphadiv[complete.cases(tmp_alphadiv),]
tmp_data <- tmp_alphadiv[tmp_alphadiv$sample_type %in% c("original_soil","suspension"),]
# select the chosen index
tmp_data <- tmp_data[,c("sample",tmp_index)]
colnames(tmp_data)[2] <- "value"
# concatenate iterations for each sample by mean
tmp_data <- tmp_data %>% 
  dplyr::group_by(sample) %>% 
  dplyr::summarise(value = mean(value))
# add treatment
tmp_data <- left_join(tmp_data,metadata_sample[,c("sample","treatment")])
# add color and order
tmp_data <- left_join(tmp_data,
                      Color_sets[Color_sets$set == "treatments",c("treatment","color","order")],
                      by = "treatment")

#####
# COMPARISON TO T0
#####
tmp_value_t0 = tmp_data$value[tmp_data$treatment =="T0"]
tmp_ttts <- unique(tmp_data$treatment)[-1]
tmp_t.test = vector()
tmp_names = vector()
for (tmp_ttt in tmp_ttts) { #tmp_ttt = tmp_ttts[1]
  # values
  tmp_value_ttt = tmp_data$value[tmp_data$treatment == tmp_ttt]
  # t-test
  tmp = t.test(tmp_value_t0,tmp_value_ttt)
  # extract p.value
  tmp_t.test[length(tmp_t.test)+1] <- tmp$p.value
  tmp_names[length(tmp_names)+1] = paste0("T0-",tmp_ttt)
}
tmp_p.adjust = tmp_t.test*length(tmp_t.test) # bonferroni correction
## letters
names(tmp_p.adjust) <- tmp_names
tmp_p.adjust
tmp_p.adjust <= 0.05

#####
# COMPARISON BETWEEN MANIPULATED SUSPENSIONS
#####
tmp_ttts <- unique(tmp_data$treatment)[3:6]
tmp_t.test = vector()
tmp_names = vector()
for (x in 1:4) {
  y = x+1
  while (y <= 4) {
    # values
    tmp_value_x <- tmp_data$value[tmp_data$treatment == tmp_ttts[x]]
    tmp_value_y <- tmp_data$value[tmp_data$treatment == tmp_ttts[y]]
    # t-test
    tmp = t.test(tmp_value_x,tmp_value_y)
    # extract p.val
    tmp_t.test[length(tmp_t.test)+1] <- tmp$p.value
    tmp_names[length(tmp_names)+1] = paste0(tmp_ttts[x],"-",tmp_ttts[y])
    # index
    y = y+1
  }
}
tmp_p.adjust = tmp_t.test*length(tmp_t.test) # bonferroni correction
## letters
names(tmp_p.adjust) <- tmp_names
tmp_p.adjust
tmp_p.adjust <= 0.05
library(multcompView)
tmp_letters <- multcompLetters(tmp_p.adjust,compare="<=",threshold=0.05,Letters=letters)
tmp_letters[["Letters"]]

```












