---
title: "3_beta_diversity"
author: "Sarah HUET"
date: "2021 M03 10"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(phyloseq)
library(ggplot2)
library(tibble)
library(dplyr)


# remove all temporary objects from your environment
rm(list = names(.GlobalEnv)[grep("tmp",names(.GlobalEnv))])

```

# Calculate distance matrix and ordination

```{r select data}

tmp_ps0 = ps_16S

tmp_sample_type = "microcosm" #  | suspension | all| modif_susp

# select sample type
if (tmp_sample_type == "microcosm") {
  tmp_ps = ps_16S_micro
  tmp_C = "C"
}
if (tmp_sample_type == "suspension") {
  tmp_ps = ps_16S_susp
  tmp_C = "susp_C"
}
if (tmp_sample_type == "modif_susp"){
  tmp_ps = ps_16S_susp
  tmp_C = "susp_C"
}
if (tmp_sample_type == "all") {
  tmp_ps = ps_16S
  tmp_C = "T0"
}

```

```{r calculate distance matrix and ordination}

# rarefy dataset
set.seed(8000)
tmp_ps = rarefy_even_depth(tmp_ps,rngseed = T)
# choose distance index
tmp_index = "wunifrac"
# calculate distances
tmp_dist = distance(tmp_ps,tmp_index)
# calculate ordination
tmp_ordin = ordinate(tmp_ps, "PCoA", distance = tmp_dist)


```

OTUs removed during rarefactions: microcosms -> 135 | susp -> 133 | all -> 2


# PCoA

```{r 2D, width = 16,height = 12 }

# select samples coordinates for ordination plot
tmp_data_2D <- tibble("sample" = as.numeric(rownames(tmp_ordin[["vectors"]])),
                      "x" = tmp_ordin[["vectors"]][,1],
                      "y" = tmp_ordin[["vectors"]][,2])
# select axis titles
tmp_x_lab = round(tmp_ordin[["values"]][["Relative_eig"]][1]*100,2)
tmp_x_lab <- paste0("Axis 1: ",tmp_x_lab," %")
tmp_y_lab = round(tmp_ordin[["values"]][["Relative_eig"]][2]*100,2)
tmp_y_lab <- paste0("Axis 2: ",tmp_y_lab," %")
# add treatment, color, order, shape, etc
tmp_data_2D <- left_join(tmp_data_2D,tmp_ps@sam_data[,c("sample","treatment")], by = "sample")
tmp_data_2D <- left_join(tmp_data_2D,Colors_treatments,by = "treatment")
tmp_data_2D <- tmp_data_2D %>% arrange(sample)
## change shape
tmp_data_2D$shape <- 21

# select data for control ellipse
tmp_data_ellipse = tmp_data_2D[tmp_data_2D$treatment == tmp_C,]


tmp_plot_title = paste0("PCOA_",tmp_index,"_",tmp_sample_type)

### plot
p = ggplot(tmp_data_2D)+
  aes(x=x,y=y) +
  geom_point(shape = tmp_data_2D$shape,
             size = 3,
             fill = tmp_data_2D$color) +
  scale_fill_manual(labels = levels(reorder(tmp_data_2D$treatment,tmp_data_2D$order)),
                    values = levels(reorder(tmp_data_2D$color,tmp_data_2D$order))) +
  scale_shape_identity() + 
  theme_bw() +
  labs(title = tmp_plot_title,x =  paste0("\n",tmp_x_lab), 
       y = paste0("\n",tmp_y_lab))+
  theme(title = element_text(face="bold", size=15),
        legend.position = "none",
        axis.title.x = element_text(face="bold", size=14), 
        axis.text.x  = element_text(vjust=0.5, size=12,angle = 0),
        axis.title.y = element_text(face="bold", size=14), 
        axis.text.y  = element_text(vjust=0.5, size=12),
        strip.text = element_text(face="bold", size=12),
        strip.background = element_rect(fill="white", colour="#333333",size=1.5))+
  guides(fill = guide_legend(override.aes = list(shape = 21,
                                                 size =12),
                             title = "Legend",
                             ncol = 3),
         size = "none")+ 
  stat_ellipse(data = tmp_data_ellipse, colour = unique(tmp_data_ellipse$color),size=2)

p


#fig_div[[paste0(tmp_sample_type,"_",tmp_index)]] <- p


```

# Data for distance barplot 

## distance to control

```{r select distances from the control}

library(reshape2)

# wrangle distance matrix into a longer dataframe
tmp_dist_matrix = melt(as.matrix(tmp_dist))
# remove self-comparisons
tmp_dist_matrix = tmp_dist_matrix[tmp_dist_matrix$Var1 != tmp_dist_matrix$Var2,]
# select sample data
tmp_sam_data = tibble("sample"=as.numeric(tmp_ps@sam_data$sample),
                      "treatment"=tmp_ps@sam_data$treatment)
# combined distance matrix with sample data
colnames(tmp_sam_data) = c("Var1", "treatment1")
tmp_data <- left_join(tmp_dist_matrix, tmp_sam_data, by = "Var1")
colnames(tmp_sam_data) = c("Var2", "treatment2")
tmp_data <- left_join(tmp_data, tmp_sam_data, by = "Var2")
# select distances from the control
tmp_data <- tmp_data[tmp_data$treatment1 == tmp_C,]

if (tmp_sample_type == "modif_susp"){
  tmp_data <- tmp_data[!tmp_data$treatment2 %in% c("T0","susp_C"),]
}

```

```{r statistic test}

library(agricolae)

# linear model
tmp_lm = lm(tmp_data$value ~ tmp_data$treatment2, na.action = na.omit )
summary(tmp_lm)
# anova
anova(tmp_lm)
tmp_aov = aov(tmp_lm)
summary(tmp_aov)
# post hoc Tukey test
tmp_comp <- HSD.test(tmp_aov,'tmp_data$treatment',alpha = 0.05,group = T)
# tibble with statistical groups
tmp_stat = tibble("treatment"=rownames(tmp_comp[["groups"]]),
                  "mean"=tmp_comp[["groups"]][["tmp_data$value"]],
                  "stat_groups"=tmp_comp[["groups"]][["groups"]])


if (tmp_sample_type %in% c("suspension","all")) {
  tmp_value_t0 = tmp_data$value[tmp_data$treatment2 =="T0"]
  tmp_ttts <- unique(tmp_data$treatment2)[-1]
  tmp_t.test = vector()
  tmp_names = vector()
  for (tmp_ttt in tmp_ttts) { #tmp_ttt = tmp_ttts[1]
    # values
    tmp_value_ttt = tmp_data$value[tmp_data$treatment2 == tmp_ttt]
    # t-test
    tmp = t.test(tmp_value_t0,tmp_value_ttt)
    # extract p.value
    tmp_t.test[length(tmp_t.test)+1] <- tmp$p.value
    tmp_names[length(tmp_names)+1] = paste0("T0-",tmp_ttt)
  }
  tmp_p.adjust = tmp_t.test*length(tmp_t.test) # bonferroni correction
  ## letters
  names(tmp_p.adjust) <- tmp_names
  tmp_p.adjust
  tmp_p.adjust <= 0.05
}


```

```{r barplot, width = 5.5,height = 6}

#library(Rmisc)

tmp_plot_data = tmp_data
tmp_plot_data = tibble("treatment" = tmp_plot_data$treatment2,
                       "sample" = tmp_plot_data$Var2,
                       "value"=tmp_plot_data$value)
tmp_plot_data = left_join(tmp_plot_data,Colors_treatments,by="treatment")
tmp_plot_stat = left_join(tmp_stat,Colors_treatments,by="treatment")
# calculate  confidence interval for comparison between the control samples
#tmp_ci = CI(tmp_data[tmp_data$treatment1 == "C" & tmp_data$treatment2 == "C","value"], ci = 0.95)

tmp_plot_title = "16S_betadiv_microcosm"
tmp_y_lim = c(0,max(tmp_data$value)*1.2)
tmp_plot_stat[,"stat_y"] = max(tmp_data$value)*1.1


### plot

p <- ggplot(tmp_plot_data, aes(x = reorder(treatment,desc(order)),
                               y = value)) +
  stat_summary(geom = "bar", fun = mean, colour = "#333333",
               linetype = 1,
               fill = levels(reorder(tmp_plot_data$color,desc(tmp_plot_data$order))))+ 
  labs(x="Treatment", y= paste0("\n",tmp_index,"Distances"), title = tmp_plot_title, color = "#333333")+
  coord_flip(ylim = tmp_y_lim)+
  #scale_y_continuous(expand = c(0,0))+
  theme_bw()+
  theme(title = element_text(face="bold", size=12),
        legend.position = "right",
        axis.title.x = element_text(face="bold", size=10), 
        axis.text.x  = element_text(vjust=0.5, size=8,angle = 0),
        axis.title.y = element_text(face="bold", size=10), 
        axis.text.y  = element_text(vjust=0.5, size=8),
        strip.text = element_text(face="bold", size=8),
        strip.background = element_rect(fill="white", colour="#333333",size=1.5))+
  stat_summary(geom = "linerange", fun.data = mean_se, width=0.5,color = "#333333" )+
  geom_text(data = tmp_plot_stat,
            mapping = aes(label = stat_groups,y = stat_y,x=treatment),
            colour="#333333", size=8,inherit.aes = F, angle = 0,
            position = position_nudge(x = -0.225))
  
p

#fig_div[[paste0(tmp_sample_type,"_",tmp_index,"_distance_tukey")]] <- p


```

# test

## multiple comparison

```{r wrangle distance matrix to df}

library(reshape2)

# wrangle distance matrix into a longer dataframe
tmp_dist_matrix = melt(as.matrix(tmp_dist))
# remove self-comparisons
tmp_dist_matrix = tmp_dist_matrix[tmp_dist_matrix$Var1 != tmp_dist_matrix$Var2,]
# select sample data
tmp_sam_data = tibble("sample"=as.numeric(tmp_ps@sam_data$sample),
                      "treatment"=tmp_ps@sam_data$treatment)
# combined distance matrix with sample data
colnames(tmp_sam_data) = c("Var1", "treatment1")
tmp_data <- left_join(tmp_dist_matrix, tmp_sam_data, by = "Var1")
colnames(tmp_sam_data) = c("Var2", "treatment2")
tmp_data <- left_join(tmp_data, tmp_sam_data, by = "Var2")
# remove T0
#tmp_data <- tmp_data[tmp_data$treatment1 != "T0" & tmp_data$treatment2 != "T0",]

```

```{r multiple comparison}
library(emmeans)

tmp_lm = lm(tmp_data$value ~tmp_data$treatment1+tmp_data$treatment2, na.action = na.omit )

# multiple comparaison
tmp_emmeans = emmeans(tmp_lm,pairwise~treatment1,adjust="tukey")
## select p value
tmp_emmeans_sum = summary(tmp_emmeans)
tmp_emmeans_sum = tmp_emmeans_sum[["contrasts"]]
## Bonferroni correction
tmp_df = tmp_emmeans_sum
#tmp_df[,"pval_adjust"] <- tmp_df$p.value * nrow(tmp_emmeans_sum)

## wrangle summary
tmp = unlist(strsplit(as.character(tmp_df$contrast)," - "))
tmp_df[,"g1"] <- tmp[seq(1,length(tmp),by=2)]
tmp_df[,"g2"] <- tmp[seq(2,length(tmp),by=2)]

    
    

```

## ellipse center

```{r TEST select distances from control ellipse center}

# get data from ggplot ellipse
tmp_data = ggplot_build(fig_div[[paste0(tmp_sample_type,"_",tmp_index)]])
tmp_data <- tmp_data$data[[2]][c("x","y")]

# calculate ellipse center
tmp_center <- MASS::cov.trob(tmp_data)$center




```

## 3d
```{r 3D}
# https://plotly-r.com/index.html
library(plotly)

# select data
tmp_data_3D <- tibble("sample" = as.numeric(rownames(tmp_ordin[["vectors"]])),
                      "x" = tmp_ordin[["vectors"]][,1],
                      "y" = tmp_ordin[["vectors"]][,2],
                      "z" = tmp_ordin[["vectors"]][,3])

tmp_x_lab = round(tmp_ordin[["values"]][["Relative_eig"]][1]*100,2)
tmp_x_lab <- paste0("Axis 1: ",tmp_x_lab," %")

tmp_y_lab = round(tmp_ordin[["values"]][["Relative_eig"]][2]*100,2)
tmp_y_lab <- paste0("Axis 2: ",tmp_y_lab," %")

tmp_z_lab = round(tmp_ordin[["values"]][["Relative_eig"]][3]*100,2)
tmp_z_lab <- paste0("Axis 3: ",tmp_z_lab," %")

# add treatment, color, order, shape, etc
tmp_data_3D <- left_join(tmp_data_3D,tmp_ps@sam_data[,c("sample","treatment")], by = "sample")
tmp_data_3D <- left_join(tmp_data_3D,Colors_treatments,by = "treatment")
tmp_data_3D <- tmp_data_3D %>% arrange(sample)

# select plot title
tmp_plot_title = paste0("PCoA 3D of ",tmp_index," distance")

### plot

p <- plot_ly(tmp_data_3D, x = ~x, y = ~y, z = ~z,
             color = reorder(tmp_data_3D$treatment,tmp_data_3D$order),
             colors = levels(reorder(tmp_data_3D$color,tmp_data_3D$order))) %>% 
  add_markers(marker = list(symbol = 'circle', 
                            sizemode = 'diameter',
                            line = list(color = '#000000',width = 2))) %>% 
  layout(title = tmp_plot_title,
                  scene = list(xaxis = list(title = tmp_x_lab),
                               yaxis = list(title = tmp_y_lab),
                               zaxis = list(title = tmp_z_lab)))

p
#gdata::mv("p",paste0("plot_pcoa_3d_",tmp_plot_name))

```

## PERMANOVA

```{r calculate similarity}

library(reshape2)

# wrangle distance matrix into a longer dataframe
tmp_dist_matrix = melt(as.matrix(tmp_dist))
# remove self-comparisons
tmp_dist_matrix = tmp_dist_matrix[tmp_dist_matrix$Var1 != tmp_dist_matrix$Var2,]
# select sample data
tmp_sam_data = tibble("sample"=tmp_ps@sam_data$sample,
                      "treatment"=tmp_ps@sam_data$treatment)
# combined distance matrix with sample data
colnames(tmp_sam_data) = c("Var1", "treatment1")
tmp_data <- left_join(tmp_dist_matrix, tmp_sam_data, by = "Var1")
colnames(tmp_sam_data) = c("Var2", "treatment2")
tmp_data <- left_join(tmp_data, tmp_sam_data, by = "Var2")
# select distances from the control
tmp_data <- tmp_data[tmp_data$treatment1 == tmp_C,]

# calculate similarity
tmp_data[,"similarity"] <- 1 -tmp_data$value

```

```{r PERMANOVA}

library(vegan)

tmp_matrix = as.matrix(tmp_dist)
str(tmp_matrix)

# check samples and treatments
tmp_ttt = tmp_ps@sam_data$treatment
tmp_ttt[1:15]
rownames(tmp_matrix)[1:15]

# adonis
tmp_adonis = adonis(formula = tmp_matrix ~ tmp_ttt, permutations = 9999)
tmp_adonis$aov.tab

library(devtools)
library(pairwiseAdonis)

tmp_adonis_pair = pairwise.adonis(tmp_matrix, factors=tmp_ttt,p.adjust.m='BH')
# subset pairs name
tmp = unlist(strsplit(tmp_adonis_pair$pairs," vs "))
tmp_adonis_pair[,"a"] <- tmp[seq(1,length(tmp),by=2)]
tmp_adonis_pair[,"b"] <- tmp[seq(2,length(tmp),by=2)]

# reorder by higher mean value
tmp_adonis_order <- tmp_data[,c("treatment2","similarity")] %>% 
  dplyr::group_by(treatment2) %>% 
  dplyr::summarise(mean = mean(similarity))
tmp_adonis_order = left_join(tmp_adonis_pair,tmp_adonis_order,
                             by=c("a"="treatment2"))

# letters
tmp = tmp_adonis_order$p.adjusted[order(tmp_adonis_order$mean,decreasing = T)]
names(tmp) <- paste0(tmp_adonis_order$a[order(tmp_adonis_order$mean,decreasing = T)],
                     "-",tmp_adonis_pair$b[order(tmp_adonis_order$mean,decreasing = T)])
tmp[1:15]
library(multcompView)
tmp_letters <- multcompLetters(tmp,compare="<=",threshold=0.05,Letters=letters)
tmp_letters[["Letters"]]

fig_div[[paste0(tmp_sample_type,"_",tmp_index,"_adonis_letters")]] <- tmp_letters[["Letters"]]

```

As we only have three replicates for the suspensions and ten for the original soil, resulting in an unbalanced design, the adonis. pairwise function doesn't work properly
https://www.researchgate.net/post/Permanova_using_Adonis2_with_unbalanced_design
https://github.com/vegandevs/vegan/issues/344


```{r PERMANOVA unbalanced}

library(vegan)

tmp_matrix = as.matrix(tmp_dist)
str(tmp_matrix)

# keep only 3 T0 samples
tmp_samples = c(131,136,140,141:143,180:191)
tmp_matrix <- tmp_matrix[rownames(tmp_matrix) %in% tmp_samples,
                         colnames(tmp_matrix) %in% tmp_samples]

# check samples and treatments
tmp_ps1 = prune_samples(sample_names(tmp_ps) %in% tmp_samples,tmp_ps)
tmp_ttt = tmp_ps1@sam_data$treatment
tmp_ttt[1:15]
rownames(tmp_matrix)[1:15]




# adonis
tmp_adonis = adonis(formula = tmp_matrix ~ tmp_ttt, permutations = 9999)
tmp_adonis$aov.tab

library(devtools)
library(pairwiseAdonis)

tmp_adonis_pair = pairwise.adonis(tmp_matrix, factors=tmp_ttt,p.adjust.m='BH')
# subset pairs name
tmp = unlist(strsplit(tmp_adonis_pair$pairs," vs "))
tmp_adonis_pair[,"a"] <- tmp[seq(1,length(tmp),by=2)]
tmp_adonis_pair[,"b"] <- tmp[seq(2,length(tmp),by=2)]

# reorder by higher mean value
tmp_adonis_order <- tmp_data[,c("treatment2","similarity")] %>% 
  dplyr::group_by(treatment2) %>% 
  dplyr::summarise(mean = mean(similarity))
tmp_adonis_order = left_join(tmp_adonis_pair,tmp_adonis_order,
                             by=c("a"="treatment2"))

# letters
tmp = tmp_adonis_order$p.adjusted[order(tmp_adonis_order$mean,decreasing = T)]
names(tmp) <- paste0(tmp_adonis_order$a[order(tmp_adonis_order$mean,decreasing = T)],
                     "-",tmp_adonis_pair$b[order(tmp_adonis_order$mean,decreasing = T)])
tmp[1:15]
library(multcompView)
tmp_letters <- multcompLetters(tmp,compare="<=",threshold=0.05,Letters=letters)
tmp_letters[["Letters"]]

fig_div[[paste0(tmp_sample_type,"_",tmp_index,"_adonis_letters")]] <- tmp_letters[["Letters"]]


```


### plot

```{r barplot similarity, width = 5.5,height = 6}

library(ggplot2)

# set limits
tmp_ylim = c(0,1)
tmp_ylim_stat = 0.95

# set plot data
tmp_plot_data = tibble("treatment"=tmp_data$treatment2,
                       "value"=tmp_data$similarity)
tmp_plot_data <- left_join(tmp_plot_data,Color_sets) # add color and order
tmp_plot_stat = tibble("treatment"=names(tmp_letters[["Letters"]]),
                       "stat_groups"=tmp_letters[["Letters"]])
tmp_plot_stat[,"stat_y"] <- tmp_ylim_stat

# plot
p <- ggplot(tmp_plot_data, aes(x = reorder(treatment,desc(order)),
                               #y = log10(value))) +
                               y = value)) +
  
  stat_summary(geom = "bar", fun = mean, colour = "#333333",
               linetype = 1,
               fill = levels(reorder(tmp_plot_data$color,desc(tmp_plot_data$order))))+ 
  labs(x="Treatment", y= "Similarity to control",
       title=paste0(tmp_sample_type,"_",tmp_index,"_similarity"),
       color = "#333333")+
  coord_flip(ylim = tmp_ylim)+
  #scale_y_log10()+
  theme_bw()+
  theme(title = element_text(face="bold", size=12),
        legend.position = "right",
        axis.title.x = element_text(face="bold", size=10), 
        axis.text.x  = element_text(vjust=0.5, size=8,angle = 0),
        axis.title.y = element_text(face="bold", size=10), 
        axis.text.y  = element_text(vjust=0.5, size=8),
        strip.text = element_text(face="bold", size=8),
        strip.background = element_rect(fill="white", colour="#333333",size=1.5))+
  stat_summary(geom = "linerange", fun.data = mean_se, width=0.5,color = "#333333" )+
  geom_text(data = tmp_plot_stat,
            #mapping = aes(label = stat_groups, y = stat_y*5, x=treatment),
            mapping = aes(label = stat_groups, y = stat_y, x=treatment),
            colour="#333333", size=8,inherit.aes = F, angle = 0,
            position = position_nudge(x = 0))
  
p

fig_div[[paste0(tmp_sample_type,"_",tmp_index,"_similarity")]] <- p



```

```{r barplot distance, width = 5.5,height = 6}

library(ggplot2)

# set limits
tmp_ylim = c(0,0.25)
tmp_ylim_stat = 0.23

# set plot data
tmp_plot_data = tibble("treatment"=tmp_data$treatment2,
                       "value"=tmp_data$value)
tmp_plot_data <- left_join(tmp_plot_data,Color_sets) # add color and order
tmp_plot_stat = tibble("treatment"=names(tmp_letters[["Letters"]]),
                       "stat_groups"=tmp_letters[["Letters"]])
tmp_plot_stat[,"stat_y"] <- tmp_ylim_stat

# plot
p <- ggplot(tmp_plot_data, aes(x = reorder(treatment,desc(order)),
                               #y = log10(value))) +
                               y = value)) +
  
  stat_summary(geom = "bar", fun = mean, colour = "#333333",
               linetype = 1,
               fill = levels(reorder(tmp_plot_data$color,desc(tmp_plot_data$order))))+ 
  labs(x="Treatment", y= "Distance to control",
       title=paste0(tmp_sample_type,"_",tmp_index,"_distance"),
       color = "#333333")+
  coord_flip(ylim = tmp_ylim)+
  #scale_y_log10()+
  theme_bw()+
  theme(title = element_text(face="bold", size=12),
        legend.position = "right",
        axis.title.x = element_text(face="bold", size=10), 
        axis.text.x  = element_text(vjust=0.5, size=8,angle = 0),
        axis.title.y = element_text(face="bold", size=10), 
        axis.text.y  = element_text(vjust=0.5, size=8),
        strip.text = element_text(face="bold", size=8),
        strip.background = element_rect(fill="white", colour="#333333",size=1.5))+
  stat_summary(geom = "linerange", fun.data = mean_se, width=0.5,color = "#333333" )+
  geom_text(data = tmp_plot_stat,
            #mapping = aes(label = stat_groups, y = stat_y*5, x=treatment),
            mapping = aes(label = stat_groups, y = stat_y, x=treatment),
            colour="#333333", size=8,inherit.aes = F, angle = 0,
            position = position_nudge(x = 0))
  
p

fig_div[[paste0(tmp_sample_type,"_",tmp_index,"_distance")]] <- p



```


