---
title: "5.2.Stat_model_diff_abund"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyr)
library(dplyr)
library(tibble)

library(lme4)
library(emmeans)


# remove all temporary objects from your environment
rm(list = names(.GlobalEnv)[grep("tmp",names(.GlobalEnv))])

```

# for OTUs

## math model

To estimate the effect of each treatment on each OTU abundance, we used a generalized linear mixed model. Considering that an OTU abundance Y, in any k replicates of any i treatment, follow a Poisson law of parameter $\Lambda$ as $Y\sim\mathcal{P}\left(\mathrm{\Lambda}\right)$, we used the following model:

$$\log{\left(\Lambda_{il}\right)}=o_{il}+\mu+\gamma_i+Z_{il},\ \ {Z_{il}}_{1\le l\le10}\mathrm{\ iid}\sim\mathcal{N}\left(0,\sigma^2\right)\ (1)$$

where:
* $i=\left\{0,\ldots,12\right\}$ represents the treatments
* $l={1,\ldots,10}$ represents the replicates
* $o$ is the offset for each sample calculated as the log of the sample read sum
* $\gamma$ is the effect of the treatments
* $Z$ is the random sampling effect modeling the data overdispersion

The analysis was performed using the glmer function of the lme4 package (version 1.1-27). Subsequently, we performed a post-hoc Tukey test with the emmeans function of the emmeans package (version 1.6.1) implementing multiple comparisons. 

## test without loop

```{r variables}

tmp_ps = ps_16S_fltr

# define model variables
## treatments
g = tmp_ps@sam_data$treatment
## offset
o = log(sample_sums(tmp_ps))
## random effect
z = tmp_ps@sam_data$sample

```

```{r model}

# select one OTU
tmp_OTU = "OTU-45"
# response variable
y = as.vector(tmp_ps@otu_table[tmp_OTU,]@.Data)

glmer1 <- glmer(y ~ -1 + g + (1 | z),family='poisson', offset = o)
glmer0 <- glmer(y ~ 1 + (1 | z),family='poisson', offset = o)



```

```{r LRT}

tmp_LRT <- 2*(logLik(glmer1)[1] - logLik(glmer0)[1])
tmp_LRT
pchisq(tmp_LRT, df=1, lower.tail=FALSE) * ntaxa(tmp_ps)

```

```{r summary}
# model summary
tmp_model_sum = summary(glmer1)$coefficients

tmp_model_sum = tibble("OTU"= tmp_OTU,
                       "treatment"=rownames(tmp_model_sum),
                       as_tibble(tmp_model_sum))

```

```{r emmeans}

# multiple comparaison
tmp_emmeans = emmeans(glmer1,pairwise~g,adjust="none")

## select p value
tmp_emmeans_sum = summary(tmp_emmeans)
tmp_emmeans_sum = tmp_emmeans_sum[["contrasts"]]
## wrangle summary
tmp_df = tmp_emmeans_sum
tmp = unlist(strsplit(as.character(tmp_df$contrast)," - "))
tmp_df[,"g1"] <- tmp[seq(1,length(tmp),by=2)]
tmp_df[,"g2"] <- tmp[seq(2,length(tmp),by=2)]
# tmp_df = tmp_df[tmp_df$g1 == "C",] # keep only comparison to control
tmp_emmeans_sum = cbind("OTU"=tmp_OTU,tmp_df)

# bonferroni p-val correction : 258 OTUs * 78 comparisons
tmp_emmeans_sum[,"pval_adjust"] <- tmp_emmeans_sum$p.value * ntaxa(tmp_ps) * nrow(tmp_emmeans_sum)
# p val significance
tmp_emmeans_sum[,"pval_signif"] <- ifelse(tmp_emmeans_sum$pval_adjust <= 0.05,"*","")
tmp_emmeans_sum[,"pval_signif"] <- ifelse(tmp_emmeans_sum$pval_adjust <= 0.01,"**",
                                                 tmp_emmeans_sum$pval_signif)
tmp_emmeans_sum[,"pval_signif"] <- ifelse(tmp_emmeans_sum$pval_adjust <= 0.001,"***",
                                                 tmp_emmeans_sum$pval_signif)

```


```{r global df}

# global df
tmp_model_sum_global = tibble()
tmp_emmeans_sum_global = tibble()

# global df
tmp_model_sum_global = rbind(tmp_model_sum_global,tmp_model_sum)
tmp_emmeans_sum_global = rbind(tmp_emmeans_sum_global,tmp_emmeans_sum)

# bonferroni p-val correction : 214 OTUs * 78 comparisons
tmp_emmeans_sum_global[,"pval_adjust"] <- tmp_emmeans_sum_global$p.value * 214 * 78
# p val significance
tmp_emmeans_sum_global[,"pval_signif"] <- ifelse(tmp_emmeans_sum_global$pval_adjust <= 0.05,"*","")
tmp_emmeans_sum_global[,"pval_signif"] <- ifelse(tmp_emmeans_sum_global$pval_adjust <= 0.01,"**",
                                                 tmp_emmeans_sum_global$pval_signif)
tmp_emmeans_sum_global[,"pval_signif"] <- ifelse(tmp_emmeans_sum_global$pval_adjust <= 0.001,"***",
                                                 tmp_emmeans_sum_global$pval_signif)


```

```{r empty summary}

# empty model summary

      tmp_ttt = c("gC","gMAC_a1d1","gMAC_a1d2","gMAC_a1d3","gMAC_a2d1","gMAC_a2d2","gMAC_a2d3",
                  "gPEB_a1d1","gPEB_a1d2","gPEB_a1d3","gPEB_a2d1","gPEB_a2d2","gPEB_a2d3")
      tmp_model_sum = tibble("OTU"= rep(tmp_OTU,length(tmp_ttt)),
                             "treatment"=tmp_ttt,
                             "Estimate"=rep(NA,length(tmp_ttt)),
                             "Std. Error"=rep(NA,length(tmp_ttt)),
                             "z value"=rep(NA,length(tmp_ttt)),
                             "Pr(>|z|)"=rep(1,length(tmp_ttt)))
      
      # empty emmeans summary
      
      
      
```

```{r remove}

rm(tmp_ps,g,o,z,tmp_model_sum_global,tmp_emmeans_sum_global)
rm(tmp_emmeans,tmp,tmp_df)
rm(y,glmer1,glmer0,tmp_LRT,tmp_LRT_pchisq,
     tmp_model_sum,tmp_emmeans_sum,tmp_OTU)

```

## loop

```{r loop}

rm(tmp_ps,g,o,z,tmp_model_sum_global,tmp_emmeans_sum_global)

tmp_ps = ps_16S_fltr

# define model variables
## treatments
g = tmp_ps@sam_data$treatment
## offset
o = log(sample_sums(tmp_ps))
## random effect
z = tmp_ps@sam_data$sample

# global df
tmp_model_sum_global = tibble()
tmp_emmeans_sum_global = tibble()




for (i in 1:ntaxa(tmp_ps)) {
  
  tmp_OTU = taxa_names(tmp_ps)[i]
  print(i)
  
  # response variable
  y = as.vector(tmp_ps@otu_table[tmp_OTU,]@.Data)
  
  glmer1 <- glmer(y ~ -1 + g + (1 | z),family='poisson', offset = o)
  glmer0 <- glmer(y ~ 1 + (1 | z),family='poisson', offset = o)
  
  # LRT
  tmp_LRT <- 2*(logLik(glmer1)[1] - logLik(glmer0)[1])
  ## calcultate Chisq pval and Bonferroni correction for 258 OTUs
  tmp_LRT_pchisq = pchisq(tmp_LRT, df=1, lower.tail=FALSE) *ntaxa(tmp_ps)

  if (tmp_LRT_pchisq <= 0.05) {
    
    # model summary
    tmp_model_sum = summary(glmer1)$coefficients
    tmp_model_sum = tibble("OTU"= tmp_OTU,
                           "treatment"=rownames(tmp_model_sum),
                           as_tibble(tmp_model_sum))
    
    # multiple comparaison
    tmp_emmeans = emmeans(glmer1,pairwise~g,adjust="none")
    ## select p value
    tmp_emmeans_sum = summary(tmp_emmeans)
    tmp_emmeans_sum = tmp_emmeans_sum[["contrasts"]]
    ## wrangle summary
    tmp_df = tmp_emmeans_sum
    tmp = unlist(strsplit(as.character(tmp_df$contrast)," - "))
    tmp_df[,"g1"] <- tmp[seq(1,length(tmp),by=2)]
    tmp_df[,"g2"] <- tmp[seq(2,length(tmp),by=2)]
    # tmp_df = tmp_df[tmp_df$g1 == "C",] # keep only comparison to control
    tmp_emmeans_sum = cbind("OTU"=tmp_OTU,tmp_df)
    
    rm(tmp_emmeans,tmp,tmp_df)
    
    } else {
      # empty model summary
      tmp_model_sum = tibble("OTU"=tmp_OTU,
                             "treatment"=NA,
                             "Estimate"=NA,
                             "Std. Error"=NA,
                             "z value"=NA,
                             "Pr(>|z|)"=1)

      
      # empty emmeans summary
      tmp_emmeans_sum = tibble("OTU"=tmp_OTU,
                               "contrast"=NA,
                               "estimate"=NA,
                               "SE"=NA,
                               "df"=NA,
                               "z.ratio"=NA,
                               "p.value"=1,
                               "g1"=NA,
                               "g2"=NA)
  }
  
  # global df
  tmp_model_sum_global = rbind(tmp_model_sum_global,tmp_model_sum)
  tmp_emmeans_sum_global = rbind(tmp_emmeans_sum_global,tmp_emmeans_sum)
  
  # remove temporary df
  rm(y,glmer1,glmer0,tmp_LRT,tmp_LRT_pchisq,
     tmp_model_sum,tmp_emmeans_sum,tmp_OTU)

}


# bonferroni p-val correction : 214 OTUs * 78 comparisons
tmp_emmeans_sum_global[,"pval_adjust"] <- tmp_emmeans_sum_global$p.value * ntaxa(tmp_ps) * 78
# p val significance
tmp_emmeans_sum_global[,"pval_signif"] <- ifelse(tmp_emmeans_sum_global$pval_adjust <= 0.05,"*","")
tmp_emmeans_sum_global[,"pval_signif"] <- ifelse(tmp_emmeans_sum_global$pval_adjust <= 0.01,"**",
                                                 tmp_emmeans_sum_global$pval_signif)
tmp_emmeans_sum_global[,"pval_signif"] <- ifelse(tmp_emmeans_sum_global$pval_adjust <= 0.001,"***",
                                                 tmp_emmeans_sum_global$pval_signif)

#stat_q2_otu_comparison_sum = tmp_emmeans_sum_global
#stat_q2_otu_model_sum = tmp_model_sum_global

```

## median filter

For the specific case where one treatment has a null abundance, we add a specific filter: for each OTU, comparisons between a treatment with a null abundance and a treatment with a median abundance higher or equal to 5, were considered significant.

```{r apply median filter for comparisons with C}

tmp_df.signif = stat_q2_otu_comparison_sum
tmp_df.signif = tmp_df.signif[tmp_df.signif$g1 == "C",] # keep only comparisons with C
tmp_df.signif <- na.omit(tmp_df.signif)
tmp_ps0 = ps_16S_fltr

# set threshold "tmp_thrld" 
tmp_thrld = 5
#calculate sum & median abundance by treatment for each OTU
tmp0 = psmelt(tmp_ps0)
tmp = tmp0[,c("OTU","treatment","Abundance")]  
tmp_abund_by_ttt = as_tibble(tmp %>% group_by(OTU,treatment) %>% summarise(sum=sum(Abundance)))
tmp_median_by_ttt = as_tibble(tmp %>% group_by(OTU,treatment) %>% summarise(median=median(Abundance)))
tmp_abund_by_ttt = left_join(tmp_abund_by_ttt,tmp_median_by_ttt)
tmp_abund_by_ttt$treatment <- as.character(tmp_abund_by_ttt$treatment)

# keep comparisons where 1 ttt have a abundance = 0 and the other ttt have a median abundance >= tmp_thrld
for (tmp_OTU in tmp_df.signif$OTU) {
  tmp_df.signif[tmp_df.signif$OTU == tmp_OTU &
                  tmp_df.signif$g2 %in% tmp_abund_by_ttt$treatment[tmp_abund_by_ttt$OTU == tmp_OTU &
                                                                    tmp_abund_by_ttt$sum == 0] &
                  tmp_df.signif$g1 %in% tmp_abund_by_ttt$treatment[tmp_abund_by_ttt$OTU == tmp_OTU &
                                                                    tmp_abund_by_ttt$median >= tmp_thrld]
                ,"pval_adjust"] <- 0
  tmp_df.signif[tmp_df.signif$OTU == tmp_OTU & 
                  tmp_df.signif$g1 %in% tmp_abund_by_ttt$treatment[tmp_abund_by_ttt$OTU == tmp_OTU &
                                                                    tmp_abund_by_ttt$sum == 0] &
                  tmp_df.signif$g2 %in% tmp_abund_by_ttt$treatment[tmp_abund_by_ttt$OTU == tmp_OTU &
                                                                    tmp_abund_by_ttt$median >= tmp_thrld]
                ,"pval_adjust"] <- 0
  }
# nb of pval <= 0.05 before and after filter
table(stat_q2_otu_comparison_sum$contrast[stat_q2_otu_comparison_sum$g1 == "C" & 
                                  stat_q2_otu_comparison_sum$pval_adjust <= 0.05])
table(tmp_df.signif$contrast[tmp_df.signif$pval_adjust <= 0.05])
# nb of OTU with a pval <= 0.05 before and after filter
tmp_otu1 = unique(stat_q2_otu_comparison_sum$OTU[stat_q2_otu_comparison_sum$g1 == "C" & 
                                                    stat_q2_otu_comparison_sum$pval_adjust <= 0.05])
tmp_otu2 = unique(tmp_df.signif$OTU[tmp_df.signif$pval_adjust <= 0.05])
# View
View(setdiff(tmp_df.signif[tmp_df.signif$pval_adjust <= 0.05,],
             stat_q2_otu_comparison_sum[stat_q2_otu_comparison_sum$g1 == "C" & 
                                  stat_q2_otu_comparison_sum$pval_adjust <= 0.05,]))
# and keep treatment with pvalue <= 0.05
stat_q2_otu_comparison_sum_w0 = tmp_df.signif

```

# results analysis

## basics

```{r global}

# select data
tmp_data = stat_q2_otu_comparison_sum

# number of significant comparisons
nrow(tmp_data[tmp_data$pval_adjust <= 0.05,])

# number of OTU with at least one significant comparison
length(unique(tmp_data$OTU[tmp_data$pval_adjust <= 0.05]))

# number of significant comparisons for each OTU
tmp_data = as.matrix(table(tmp_data$OTU[tmp_data$pval_adjust <= 0.05]))
tmp_data <- tibble("OTU"=rownames(tmp_data),
                   "nb_signif_comp"=tmp_data)

p <- ggplot(tmp_data,aes(tmp_data$nb_signif_comp))+
  geom_histogram(binwidth = 1)+
  theme_bw()+
  labs(title = "Number of significant comparisons by OTUs",
       x= "Number of significant comparisons",
       y = "Number of OTUs")
p

```

## comparison with control

```{r comparison with control}

# select data
tmp_comp = stat_q2_otu_comparison_sum
tmp_comp = tmp_comp[complete.cases(tmp_comp),] # remove OTUs for which model didnt work
tmp_comp = tmp_comp[tmp_comp$g1 == "C",]
tmp_comp$pval_adjust <- tmp_comp$p.value *258 *12

# number of significant comparisons
nrow(tmp_comp[tmp_comp$pval_adjust <= 0.05,])

# number of OTU with at least one significant comparison
length(unique(tmp_comp$OTU[tmp_comp$pval_adjust <= 0.05]))

# number of OTU C.Tup and C.Tdown
## estimate < 0 => superior in control than in ttt
length(unique(tmp_comp$OTU[tmp_comp$pval_adjust <= 0.05 & tmp_comp$estimate < 0]))
length(unique(tmp_comp$OTU[tmp_comp$pval_adjust <= 0.05 & tmp_comp$estimate > 0]))

```

```{r taxo impacted OTUs}

# number of OTU C.Tup and C.Tdown
## estimate < 0 => superior in control than in ttt
tmp_otu_up = unique(tmp_comp$OTU[tmp_comp$pval_adjust <= 0.05 & tmp_comp$estimate < 0])
tmp_otu_up = tmp_otu_up[!is.na(tmp_otu_up)]
tmp_otu_down = unique(tmp_comp$OTU[tmp_comp$pval_adjust <= 0.05 & tmp_comp$estimate > 0])
tmp_otu_down = tmp_otu_down[!is.na(tmp_otu_down)]

tmp_df = ps_16S@tax_table[c(tmp_otu_up)]@.Data
table(tmp_df[,3])

tmp_df = ps_16S@tax_table[c(tmp_otu_down)]@.Data
table(tmp_df[,3])

# common OTUs between up and down
intersect(tmp_otu_up,tmp_otu_down)
tmp_df = ps_16S@tax_table[c(intersect(tmp_otu_up,tmp_otu_down))]@.Data
table(tmp_df[,3])

```


```{r by treatment}

# number of significant comparisons for each treatment
tmp_total = as.matrix(table(tmp_comp$g2[tmp_comp$pval_adjust <= 0.05]))
tmp_C.Td = as.matrix(table(tmp_comp$g2[tmp_comp$pval_adjust <= 0.05 & tmp_comp$estimate >0]))
tmp_C.Tu = as.matrix(table(tmp_comp$g2[tmp_comp$pval_adjust <= 0.05 & tmp_comp$estimate <0]))

tmp_data <- tibble("treatment"=rownames(tmp_C.Td),
                   "total"= tmp_total,
                   "C>T"=tmp_C.Td,
                   "C<T"=tmp_C.Tu)

tmp_data$`C>T` = - tmp_data$`C>T`

tmp_data <- pivot_longer(tmp_data, cols = !treatment, names_to = "comp",values_to = "count")
tmp_data <- left_join(tmp_data,Colors_treatments,by = "treatment")

p <- ggplot(data = tmp_data,
            aes(x = reorder(treatment,desc(order)), y = count)) + 
  geom_col(colour = "#333333",linetype = 1, fill = reorder(tmp_data$color,desc(tmp_data$order))) +
  coord_flip()+
  labs(x="Treatment", y= "Differentially abundant OTU count", 
       title = "Number of OTU differentially abundant in treatment compared to control",
       color = "#333333")+
  theme_bw()+
  theme(title = element_text(face="bold", size=15),
        legend.position = "right",
        axis.title.x = element_text(face="bold", size=14), 
        axis.text.x  = element_text(vjust=0.5, size=12,angle = 0),
        axis.title.y = element_text(face="bold", size=14), 
        axis.text.y  = element_text(vjust=0.5, size=12),
        strip.text = element_text(face="bold", size=12),
        strip.background = element_rect(fill="white", colour="#333333",size=1.5))+
  geom_hline(yintercept = 0,color = "#333333")+
  geom_text(data = tmp_data[tmp_data$comp =="total",],
            mapping = aes(label = tmp_data$count[tmp_data$comp =="total"],
                          y = 0.5,
                          x = tmp_data$treatment[tmp_data$comp =="total"]),
            colour="#333333", size=10,inherit.aes = F, angle = 0,
            position = position_nudge(x = 0))

p

```

```{r by OTU}

# number of significant comparisons for each OTU
tmp_data = as.matrix(table(tmp_comp$OTU[tmp_comp$pval_adjust <= 0.05]))
tmp_data <- tibble("OTU"=rownames(tmp_data),
                   "nb_signif_comp"=tmp_data)

p <- ggplot(tmp_data,aes(tmp_data$nb_signif_comp))+
  geom_histogram(binwidth = 1, breaks = c(0:max(tmp_data$nb_signif_comp)))+
  theme_bw()+
  labs(title = "Number of significant comparisons with the control by OTUs",
       x= "Number of significant comparisons with the control",
       y = "Number of OTUs")
p

```

### impacted OTUs in control (suspension)

```{r}

# select data
tmp_comp = stat_q2_otu_comparison_sum
tmp_comp = tmp_comp[complete.cases(tmp_comp),] # remove OTUs for which model didnt work
tmp_comp = tmp_comp[tmp_comp$g1 == "C",]
tmp_comp$pval_adjust <- tmp_comp$p.value *258 *12
tmp_comp = tmp_comp[tmp_comp$pval_adjust <= 0.05,]

# impacted otus
tmp_OTUs_impacted = unique(tmp_comp$OTU)

# OTUs in control susp
tmp_ps = ps_16S_susp
tmp_ps <- prune_samples(tmp_ps@sam_data$treatment == "susp_C",tmp_ps)
tmp_ps <- prune_taxa(taxa_sums(tmp_ps) >0, tmp_ps)
tmp_OTUs_susp_C = taxa_names(tmp_ps)

# OTUs impacted not in susp C
setdiff(tmp_OTUs_impacted,tmp_OTUs_susp_C)

view(tmp_comp[tmp_comp$OTU %in% setdiff(tmp_OTUs_impacted,tmp_OTUs_susp_C),])

# OTUs in control
tmp_ps = ps_16S
tmp_ps <- prune_samples(tmp_ps@sam_data$treatment == "C",tmp_ps)
tmp_ps <- prune_taxa(taxa_sums(tmp_ps) >0, tmp_ps)
tmp_OTUs_C = taxa_names(tmp_ps)

# OTUs impacted not in C
setdiff(tmp_OTUs_impacted,tmp_OTUs_C)



```


### impacted in corresponding suspensions
How many OTU impacted by the coalescence are present in the suspensions ?

```{r without loop}

tmp_susp = c("susp_MAC_a1","susp_MAC_a2","susp_PEB_a1","susp_PEB_a2")
tmp_ttt = list(c("MAC_a1d1","MAC_a1d2","MAC_a1d3"),c("MAC_a2d1","MAC_a2d2","MAC_a2d3"),
               c("PEB_a1d1","PEB_a1d2","PEB_a1d3"),c("PEB_a2d1","PEB_a2d2","PEB_a2d3"))

i = 1
  
# select comparison with control
tmp_data = stat_q2_otu_comparison_sum
tmp_data = tmp_data[complete.cases(tmp_data),] # remove OTUs for which model didnt work
tmp_data = tmp_data[tmp_data$g1 == "C",]
tmp_data$pval_adjust <- tmp_data$p.value *258 *12

# select treatments and signif comp up
tmp_data = tmp_data[tmp_data$g2 %in% tmp_ttt[[i]] & # select comp
                      tmp_data$pval_adjust <= 0.05 & # signif comp
                      tmp_data$estimate <0,] # OTUs up < | down >

# select OTUs up
tmp_OTUs_up = unique(tmp_data$OTU)

# select OTU in suspension
tmp_ps = ps_16S_susp
tmp_ps = prune_samples(tmp_ps@sam_data$treatment == tmp_susp[i],tmp_ps)
tmp_ps = prune_taxa(taxa_sums(tmp_ps) >0, tmp_ps)
tmp_OTUs_susp = taxa_names(tmp_ps)

# compare OTUs present
tmp_OTUs_pres = intersect(tmp_OTUs_up,tmp_OTUs_susp)
# compare OTUs absent
tmp_OTUs_abs = setdiff(tmp_OTUs_up,tmp_OTUs_susp)

# global df
tmp_df_pres = tibble("susp"=rep(tmp_susp[i],length(tmp_OTUs_pres)),
                     "OTU"=tmp_OTUs_pres)
tmp_df_abs = tibble("susp"=rep(tmp_susp[i],length(tmp_OTUs_abs)),
                      "OTU"=tmp_OTUs_abs)


```

```{r all suspensions, up}

tmp_susp = c("susp_MAC_a1","susp_MAC_a2","susp_PEB_a1","susp_PEB_a2")
tmp_ttt = list(c("MAC_a1d1","MAC_a1d2","MAC_a1d3"),c("MAC_a2d1","MAC_a2d2","MAC_a2d3"),
               c("PEB_a1d1","PEB_a1d2","PEB_a1d3"),c("PEB_a2d1","PEB_a2d2","PEB_a2d3"))

tmp_global_up_pres = tibble()
tmp_global_up_abs = tibble()


for (i in 1:4) {
  
  # select comparison with control
  tmp_data = stat_q2_otu_comparison_sum
  tmp_data = tmp_data[complete.cases(tmp_data),] # remove OTUs for which model didnt work
  tmp_data = tmp_data[tmp_data$g1 == "C",]
  tmp_data$pval_adjust <- tmp_data$p.value *258 *12
  
  # select treatments and signif comp up
  tmp_data = tmp_data[tmp_data$g2 %in% tmp_ttt[[i]] & # select comp
                        tmp_data$pval_adjust <= 0.05 & # signif comp
                        tmp_data$estimate <0,] # OTUs up <
  
  # select OTUs up
  tmp_OTUs_up = unique(tmp_data$OTU)
  
  # select OTU in suspension
  tmp_ps = ps_16S_susp
  tmp_ps = prune_samples(tmp_ps@sam_data$treatment == tmp_susp[i],tmp_ps)
  tmp_ps = prune_taxa(taxa_sums(tmp_ps) >0, tmp_ps)
  tmp_OTUs_susp = taxa_names(tmp_ps)
  
  # compare OTUs present
  tmp_OTUs_pres = intersect(tmp_OTUs_up,tmp_OTUs_susp)
  # compare OTUs absent
  tmp_OTUs_abs = setdiff(tmp_OTUs_up,tmp_OTUs_susp)

  # global df
  tmp_df_pres = tibble("susp"=rep(tmp_susp[i],length(tmp_OTUs_pres)),
                       "OTU"=tmp_OTUs_pres)
  tmp_df_abs = tibble("susp"=rep(tmp_susp[i],length(tmp_OTUs_abs)),
                      "OTU"=tmp_OTUs_abs)
  
  tmp_global_up_pres = rbind(tmp_global_up_pres,tmp_df_pres)
  tmp_global_up_abs = rbind(tmp_global_up_abs,tmp_df_abs)
  
}

unique(tmp_global_up_pres$OTU)
unique(tmp_global_up_abs$OTU)

# taxo
tmp_df = ps_16S@tax_table[c(unique(tmp_global_up_pres$OTU))]@.Data
table(tmp_df[,3])

tmp_df = ps_16S@tax_table[c(unique(tmp_global_up_abs$OTU))]@.Data
table(tmp_df[,3])

```

```{r all suspensions, down}

tmp_susp = c("susp_MAC_a1","susp_MAC_a2","susp_PEB_a1","susp_PEB_a2")
tmp_ttt = list(c("MAC_a1d1","MAC_a1d2","MAC_a1d3"),c("MAC_a2d1","MAC_a2d2","MAC_a2d3"),
               c("PEB_a1d1","PEB_a1d2","PEB_a1d3"),c("PEB_a2d1","PEB_a2d2","PEB_a2d3"))

tmp_global_down_pres = tibble()
tmp_global_down_abs = tibble()

for (i in 1:4) {
  
  # select comparison with control
  tmp_data = stat_q2_otu_comparison_sum
  tmp_data = tmp_data[complete.cases(tmp_data),] # remove OTUs for which model didnt work
  tmp_data = tmp_data[tmp_data$g1 == "C",]
  tmp_data$pval_adjust <- tmp_data$p.value *258 *12
  
  # select treatments and signif comp up
  tmp_data = tmp_data[tmp_data$g2 %in% tmp_ttt[[i]] & # select comp
                        tmp_data$pval_adjust <= 0.05 & # signif comp
                        tmp_data$estimate >0,] # OTUs down > 
  
  # select OTUs down
  tmp_OTUs_down = unique(tmp_data$OTU)
  
  # select OTU in suspension
  tmp_ps = ps_16S_susp
  tmp_ps = prune_samples(tmp_ps@sam_data$treatment == tmp_susp[i],tmp_ps)
  tmp_ps = prune_taxa(taxa_sums(tmp_ps) >0, tmp_ps)
  tmp_OTUs_susp = taxa_names(tmp_ps)
  
  # compare OTUs present
  tmp_OTUs_pres = intersect(tmp_OTUs_down,tmp_OTUs_susp)
  # compare OTUs absent
  tmp_OTUs_abs = setdiff(tmp_OTUs_down,tmp_OTUs_susp)

  # global df
  tmp_df_pres = tibble("susp"=rep(tmp_susp[i],length(tmp_OTUs_pres)),
                       "OTU"=tmp_OTUs_pres)
  tmp_df_abs = tibble("susp"=rep(tmp_susp[i],length(tmp_OTUs_abs)),
                      "OTU"=tmp_OTUs_abs)
  
  tmp_global_down_pres = rbind(tmp_global_down_pres,tmp_df_pres)
  tmp_global_down_abs = rbind(tmp_global_down_abs,tmp_df_abs)
  
}

unique(tmp_global_down_pres$OTU)
unique(tmp_global_down_abs$OTU)

```

```{r percent impacted / present in susp}

tmp_df = tibble("OTU"=taxa_names(ps_16S_fltr))

tmp_OTUs = tmp_otu_up
tmp_df[,"up"] <- ifelse(tmp_df$OTU %in% tmp_OTUs,1,0)

tmp_OTUs = unique(tmp_df_final_up$OTU)
tmp_df[,"up_in_susp"] <- ifelse(tmp_df$OTU %in% tmp_OTUs,1,0)

tmp_OTUs = tmp_otu_down
tmp_df[,"down"] <- ifelse(tmp_df$OTU %in% tmp_OTUs,1,0)

tmp_OTUs = unique(tmp_df_final_down$OTU)
tmp_df[,"down_in_susp"] <- ifelse(tmp_df$OTU %in% tmp_OTUs,1,0)

tmp <- apply(tmp_df[,-1], 1, FUN=function(x) sum(x))
tmp_OTUs <- tmp_df$OTU[which(tmp>0)]

# nb of OTUs impacted by the coalescence
length(tmp_OTUs)/258*100

# nb of OTUs impacted by the coalescence that were in modified suspensions
length(unique(c(tmp_df$OTU[tmp_df$up_in_susp >0],
                tmp_df$OTU[tmp_df$down_in_susp >0]))) /length(tmp_OTUs)*100

```

```{r percent impacted / absent in susp}

# select signif comp with control
tmp_comp = stat_q2_otu_comparison_sum
tmp_comp = tmp_comp[complete.cases(tmp_comp),] # remove OTUs for which model didnt work
tmp_comp = tmp_comp[tmp_comp$g1 == "C",]
tmp_comp$pval_adjust <- tmp_comp$p.value *258 *12
tmp_comp = tmp_comp[tmp_comp$pval_adjust <= 0.05,]

# number of OTU with at least one significant comparison
tmp_otus_impacted = unique(tmp_comp$OTU)

# number of OTUs impacted present in suspension
tmp_OTUs_pres = unique(c(tmp_global_up_pres$OTU,tmp_global_down_pres$OTU))
## OTU up present
tmp_otus_up_pres = unique(tmp_global_up_pres$OTU)
## OTU down present
tmp_otus_down_pres = unique(tmp_global_down_pres$OTU)

# number of OTUs impacted absent from suspension
tmp_OTUs_abs = unique(c(tmp_global_up_abs$OTU,tmp_global_down_abs$OTU))
## OTU up absent
tmp_otus_up_abs = unique(tmp_global_up_abs$OTU)
## OTU down absent
tmp_otus_down_abs = unique(tmp_global_down_abs$OTU)





tmp_df = tibble("OTU"=taxa_names(ps_16S_fltr))

tmp_OTUs = tmp_otu_up
tmp_df[,"up"] <- ifelse(tmp_df$OTU %in% tmp_OTUs,1,0)

tmp_OTUs = unique(tmp_df_final_up$OTU)
tmp_df[,"up_in_susp"] <- ifelse(tmp_df$OTU %in% tmp_OTUs,1,0)

tmp_OTUs = tmp_otu_down
tmp_df[,"down"] <- ifelse(tmp_df$OTU %in% tmp_OTUs,1,0)

tmp_OTUs = unique(tmp_df_final_down$OTU)
tmp_df[,"down_in_susp"] <- ifelse(tmp_df$OTU %in% tmp_OTUs,1,0)

tmp <- apply(tmp_df[,-1], 1, FUN=function(x) sum(x))
tmp_OTUs <- tmp_df$OTU[which(tmp>0)]

# nb of OTUs impacted by the coalescence
length(tmp_OTUs)/258*100

# nb of OTUs impacted by the coalescence that were in modified suspensions
length(unique(c(tmp_df$OTU[tmp_df$up_in_susp >0],
                tmp_df$OTU[tmp_df$down_in_susp >0]))) /length(tmp_OTUs)*100

```


# plot OTU abundance

```{r plot 1 OTU abundance}

tmp_OTU = "OTU-47"

tmp_ps = ps_16S_fltr
## transform to relative abundance 
tmp_ps <- transform_sample_counts(tmp_ps,function(x) x / sum(x))
## prune taxa
tmp_ps = prune_taxa(taxa_names(tmp_ps)==tmp_OTU,tmp_ps)

tmp_data = psmelt(tmp_ps)
# treatment colors
tmp_colors <- Color_sets[Color_sets$set == "treatments",c("treatment","color","order")]  
tmp_colors <- tmp_colors[tmp_colors$treatment %in% tmp_data$treatment,]


p <- ggplot(tmp_data,aes(tmp_data$treatment,tmp_data$Abundance,fill = treatment))+
  geom_boxplot()+
  geom_jitter(shape = 21,size=2, alpha=0.5) +
  labs(title = tmp_OTU,y="Relative abundance (among non filtered OTUs)",x="Treatments")+
  scale_fill_manual(values = tmp_colors$color) +
  #scale_color_manual(values = tmp_colors$color) +
  #labs(title = tmp_OTU,y="Raw abundance (log10 scale)",x="Treatments")+
  #scale_y_log10()+
  geom_hline(yintercept = 0.25/100)+
  geom_hline(yintercept = 0.1/100)+
  theme(axis.text.x  = element_text(angle = 90))
  
p

```


```{r plot several OTU abundance}

tmp_OTUs = c(2352,17,12250,5544,17138)
tmp_OTUs <- paste0("OTU-",tmp_OTUs)

#tmp_OTUs = tmp_otus_down_abs

tmp_ps = ps_16S_fltr
tmp_ps <- prune_taxa(taxa_names(tmp_ps) %in% tmp_OTUs,tmp_ps)

tmp_plot_data = psmelt(tmp_ps)


p <- ggplot(tmp_plot_data,aes(tmp_plot_data$treatment,tmp_plot_data$Abundance))+
  geom_boxplot()+
  geom_jitter(color="red", size=0.5, alpha=0.9) +
  labs(title = "Node OTUs",y="Abundance",x="Treatments")+
#  labs(title = tmp_OTU,y="Raw abundance (log10 scale)",x="Treatments")+
  scale_y_log10()+
  geom_hline(yintercept = 0.25/100)+
  geom_hline(yintercept = 0.1/100)+
  facet_wrap(vars(OTU))+
  theme(axis.text.x  = element_text(angle = 90))

p

```










