---
title: "7_comp_susp_microcosm"
author: "Sarah HUET"
date: '2023-03-27'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(phyloseq)
library(PLNmodels)
library(tibble)
library(stringr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggplot2)

library(gdata)

# load data
load("E:/Research/Thesis/EMFEED/BISCAL_xp2_coal/R_xp2_coal/.RData")
# remove all temporary objects from your environment
rm(list = names(.GlobalEnv)[grep("tmp",names(.GlobalEnv))])

```

# all suspensions

I want to compare OTU's behavior in suspenions and microcosm

```{r OTUs in susp}

# OTUs in susp
tmp_ps = ps_16S_susp
## remove T0 samples
tmp_ps <- prune_samples(tmp_ps@sam_data$treatment != "T0",tmp_ps)
tmp_ps <- prune_taxa(taxa_sums(tmp_ps) >0,tmp_ps)
## melt ps
tmp_ps_melt = psmelt(tmp_ps)

# suspension names
tmp_name_susp = c("all","susp_C","susp_manip")

# list of OTUs
tmp_OTUs_susp = list()
## OTUs all susp
tmp_OTUs_susp[[tmp_name_susp[1]]] = taxa_names(tmp_ps)
## OTUs in susp_C
tmp_OTUs_susp[[tmp_name_susp[2]]] = unique(tmp_ps_melt$OTU[tmp_ps_melt$treatment == "susp_C" 
                                         & tmp_ps_melt$Abundance > 0])
## OTUs in manipulated suspensions
tmp_OTUs_susp[[tmp_name_susp[3]]] = unique(tmp_ps_melt$OTU[tmp_ps_melt$treatment != "susp_C"
                                             & tmp_ps_melt$Abundance > 0])

```

Which OTUs in suspensions are also detected in microcosms?

```{r OTUs both in susp and microcosms}

tmp_OTUs_microcosms = taxa_names(ps_16S_micro)
tmp_OTUs_microcosms_fltr = taxa_names(ps_16S_fltr)

for (tmp_i in tmp_name_susp) {
  print(tmp_i)
  print(length(intersect(tmp_OTUs_susp[[tmp_i]],tmp_OTUs_microcosms)))
  print(length(intersect(tmp_OTUs_susp[[tmp_i]],tmp_OTUs_microcosms_fltr)))

}

# which OTUs both in susp_manip & most abund but not in susp_C
setdiff(intersect(tmp_OTUs_susp$susp_manip,tmp_OTUs_microcosms_fltr),tmp_OTUs_susp$susp_C)

```

How are OTUs detected in manipulated suspensions affected by coalescence?

```{r effect of coal}

# GLMM results for OTUs in susp
tmp_data = list()

for (tmp_i in tmp_name_susp) {
  tmp_data[[tmp_i]] = stat_q2_otu_comparison_sum_w0 %>%
    dplyr::filter(OTU %in% tmp_OTUs_susp[[tmp_i]] & pval_adjust <= 0.05)
  
  ## nb OTUs impacted
  print(length(unique(tmp_data[[tmp_i]]$OTU)))
  
  ## nb OTUs positively impacted /!\ estimate < 0
  print(length(unique(tmp_data[[tmp_i]]$OTU[tmp_data[[tmp_i]]$estimate <0])))
  ## nb OTUs negatively impacted /!\ estimate > 0
  print(length(unique(tmp_data[[tmp_i]]$OTU[tmp_data[[tmp_i]]$estimate >0])))
  
}


```

How community properties influence OTUs abundance?

```{r effect of properties}

# GLMM results for OTUs in susp
tmp_data = list()

for (tmp_i in tmp_name_susp) {
  print(tmp_i)
  tmp_data[[tmp_i]] = stat_q1_otu %>%
    dplyr::filter(OTU %in% tmp_OTUs_susp[[tmp_i]] & pval_adjust <= 0.05)
  
  ## nb OTUs affected by properties
  print(length(unique(tmp_data[[tmp_i]]$OTU)))
  
  for (tmp_j in unique(tmp_data[[tmp_i]]$var)) {
    print(tmp_j)
    ## nb OTUs affected by var
    print(length(unique(tmp_data[[tmp_i]]$OTU[tmp_data[[tmp_i]]$var == tmp_j])))
  }
  
}

```

How many OTUs in susp are connected in network?

```{r}

# PLN results for OTUs in susp
tmp_data = list()

for (tmp_i in tmp_name_susp) {
  print(tmp_i)
  tmp_data[[tmp_i]] = PLN_model.StARS_xp2_edges %>%
    dplyr::filter(source %in% tmp_OTUs_susp[[tmp_i]] | 
                    target %in% tmp_OTUs_susp[[tmp_i]])
  
  ## nb OTUs connected in network
  print(table(tmp_OTUs_susp[[tmp_i]] %in% c(PLN_model.StARS_xp2_edges$source,PLN_model.StARS_xp2_edges$target)))

  
}
```

# Most abund OTUs in  susp

I will calculate the relative abundance among the 258 most abundant OTUs as I will only display and talk about these ones. If I filter the OTUs after calculating the relative abundance, the sum of the rel ab is <1 in each suspension.

## rel ab in susp

```{r LRT a la main}

tmp_ps0 = ps_16S_susp
## remove T0 samples
tmp_ps0 <- prune_samples(tmp_ps0@sam_data$treatment != "T0",tmp_ps0)
tmp_ps0 <- prune_taxa(taxa_sums(tmp_ps0) >0,tmp_ps0)

## subset taxa
tmp_ps0 = prune_taxa(taxa_names(tmp_ps0) %in% taxa_names(ps_16S_fltr),tmp_ps0)

# loop
tmp_loop_df = tibble()
tmp_name_susp = c("MAC_a1","MAC_a2","PEB_a1","PEB_a2")
for (tmp_i in taxa_names(tmp_ps0)) {
  tmp_OTU = tmp_i
  for (tmp_j in paste0("susp_",tmp_name_susp)) {
    tmp_ttt1 = "susp_C"
    tmp_ttt2 = tmp_j
    
    # subset samples
    tmp_ps = prune_samples(tmp_ps0@sam_data$treatment %in% c(tmp_ttt1,tmp_ttt2),tmp_ps0)
    
    #  treatment
    tmp_ttt = unlist(tmp_ps@sam_data$treatment)
    names(tmp_ttt) <- unlist(tmp_ps@sam_data$sample)

    # response variable
    tmp_y = unlist(tmp_ps@otu_table@.Data[tmp_OTU,])
    tmp_relab = tmp_y / sample_sums(tmp_ps)

    # a la main
    tmp_logLik1 <- sum(dpois(tmp_y[tmp_ttt==tmp_ttt1],
                            mean(tmp_y[tmp_ttt==tmp_ttt1]), log=TRUE)) +
     sum(dpois(tmp_y[tmp_ttt==tmp_ttt2],
               mean(tmp_y[tmp_ttt==tmp_ttt2]), log=TRUE))
    tmp_logLik0 <- sum(dpois(tmp_y, mean(tmp_y), log=TRUE)) 
    tmp_LRT <- 2*(tmp_logLik1 - tmp_logLik0)
    tmp_LRT
    tmp_pchisq = pchisq(tmp_LRT, df=1, lower.tail=FALSE)

    
    tmp <- tibble("OTU"=tmp_OTU,
                  "susp_C_relab"=mean(tmp_relab[tmp_ttt==tmp_ttt1]),
                  "manip_susp"=tmp_j,
                  "manip_susp_relab"=mean(tmp_relab[tmp_ttt==tmp_ttt2]),
                  "LRT"=tmp_LRT,
                  "pchisq"=tmp_pchisq,
                  "p_adjust"=tmp_pchisq*length(taxa_names(tmp_ps0))*4)
    tmp[,"ratio"] <- ifelse(tmp$manip_susp_relab == 0,0,
                            tmp$manip_susp_relab/(tmp$susp_C_relab + tmp$manip_susp_relab))
    # add a ratio threshold to consider higher or lower
    tmp[,"ratio_cat"] <- ifelse(tmp$p_adjust > 0.05,"n.s.",
                                ifelse(tmp$ratio <=0.4,"lower_in_manip",
                                       ifelse(tmp$ratio >=0.6,"higher_in_manip",
                                              "n.s.")))

    tmp_loop_df <- rbind(tmp_loop_df,tmp)

  }
  
}

#stat_susp_OTU = tmp_loop_df

# OTUs in manipulated susp and most abund in microcosms
tmp_OTUs_56 = intersect(taxa_names(ps_16S_susp_manip),taxa_names(ps_16S_fltr))
# which OTU higher in manip susp
tmp_OTUs_higher <- unique(tmp_loop_df$OTU[tmp_loop_df$ratio_cat =="higher_in_manip"])



```


```{r plot rel ab points}

# PLOT REL AB IN SUSP
tmp_name_susp = c("MAC_a1","MAC_a2","PEB_a1","PEB_a2")

for (tmp_i in paste0("susp_",tmp_name_susp)) {
  print(
    stat_susp_OTU %>%
      filter(manip_susp %in% tmp_i) %>%
      ggplot() +
      aes(x = susp_C_relab, y = manip_susp_relab, fill = ratio_cat,size=abs(ratio-0.5)) +
      geom_point(shape = "circle filled") +
      geom_abline(intercept = 0) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      scale_x_continuous(trans = "asn") +
      scale_y_continuous(trans = "asn") +
      scale_fill_manual(values = c("#FF3366","#3399FF","#cccccc")) +
      scale_size_continuous(limits = c(0, 0.5)) +
      theme_light() +
      labs(title = tmp_i)
    )
}

# rank curve
for (tmp_i in paste0("susp_",tmp_name_susp)) {
  p <- stat_susp_OTU %>%
      filter(manip_susp %in% tmp_i) %>%
      ggplot() +
      geom_col(aes(x=reorder(OTU,desc(susp_C_relab)),y=susp_C_relab),fill = "#cccccc")+
      geom_col(aes(x=reorder(OTU,desc(susp_C_relab)),y=manip_susp_relab,fill = ratio_cat))+
      geom_vline(xintercept = seq(10.5,236,10),lwd=0.1, color = "#cccccc") +
      scale_y_continuous(trans = "asn") +
      scale_fill_manual(values = c("#FF3366","#3399FF","#333333"))+
      labs(title = tmp_i) +
      theme_light() +
      theme(axis.text.x  = element_text(vjust=0.5, size=5,angle = 90),
        panel.grid.major.x = element_blank())
  print(p)
  
  library(ggedit)
  # export svg
  ggsave(plot = p,
         filename = paste0("C:/Users/srhuet/Downloads/tmp_fig",tmp_i,".svg"),
         width = 200,height = 100,units = "mm")
}

```

### itol

```{r piechart 1 ratio_cat}

tmp = tibble("OTU" = unique(stat_susp_OTU$OTU))
tmp[,"ratio_cat"] <- ifelse(tmp$OTU %in%
                              unique(stat_susp_OTU$OTU[stat_susp_OTU$ratio_cat == "higher_in_manip"]),
                            "higher_in_manip",
                            ifelse(tmp$OTU %in%
                                     unique(stat_susp_OTU$OTU[stat_susp_OTU$ratio_cat == "lower_in_manip"]),
                                   "lower_in_manip","n.s."))
tmp[,"values"] <- 1
tmp <- pivot_wider(tmp, names_from = ratio_cat,values_from = values,values_fill = 0)

tmp_itol = tibble("OTU"= taxa_names(ps_16S_fltr),
                  "external"=-1,
                  "radius"=50) %>%
  left_join(.,tmp,by="OTU")
tmp_itol[is.na(tmp_itol)] <- 0

```


```{r itol piechart 4 susp}

tmp_itol = tibble("OTU"= taxa_names(ps_16S_fltr),
                  "external"=-1,
                  "radius"=50)

tmp = stat_susp_OTU %>%
  dplyr::group_by(OTU,ratio_cat) %>%
  dplyr::summarise(piechart = n()) %>%
  pivot_wider(.,names_from = ratio_cat,values_from = piechart,values_fill = 0)

tmp_itol <- left_join(tmp_itol,tmp)
tmp_itol[is.na(tmp_itol)] <- 0

# extract data
#write.csv(tmp_itol,"iTOL/tmp_itol_tree_data.csv",row.names = F,quote = F)
# COLOR_MIN,#3399FF / COLOR_MAX,#FF3366

```


```{r itol binary}

# select data
tmp = stat_susp_OTU

# insert data in 258 most abund OTUs
## data for n.s.
tmp_data = tibble("OTU"= taxa_names(ps_16S_fltr))
tmp_data[,"susp_C"] <- ifelse(tmp_data$OTU %in% tmp$OTU[tmp$susp_C_relab >0],1,-1)
tmp_data[,"susp_MAC_a1"] <- ifelse(tmp_data$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_MAC_a1" &
                                                tmp$manip_susp_relab >0],0,-1)
tmp_data[,"susp_MAC_a2"] <- ifelse(tmp_data$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_MAC_a2" &
                                                tmp$manip_susp_relab >0],0,-1)
tmp_data[,"susp_PEB_a1"] <- ifelse(tmp_data$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_PEB_a1" &
                                                tmp$manip_susp_relab >0],0,-1)
tmp_data[,"susp_PEB_a2"] <- ifelse(tmp_data$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_PEB_a2" &
                                                tmp$manip_susp_relab >0],0,-1)
## data for higher_in_manip
tmp_data_h = tibble("OTU"= taxa_names(ps_16S_fltr))
tmp_data_h[,"susp_C"] <- -1
tmp_data_h[,"susp_MAC_a1"] <- ifelse(tmp_data_h$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_MAC_a1" &
                                                tmp$manip_susp_relab >0 &
                                                tmp$ratio_cat == "higher_in_manip"],1,-1)
tmp_data_h[,"susp_MAC_a2"] <- ifelse(tmp_data_h$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_MAC_a2" &
                                                tmp$manip_susp_relab >0 &
                                                tmp$ratio_cat == "higher_in_manip"],1,-1)
tmp_data_h[,"susp_PEB_a1"] <- ifelse(tmp_data_h$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_PEB_a1" &
                                                tmp$manip_susp_relab >0 &
                                                tmp$ratio_cat == "higher_in_manip"],1,-1)
tmp_data_h[,"susp_PEB_a2"] <- ifelse(tmp_data_h$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_PEB_a2" &
                                                tmp$manip_susp_relab >0 &
                                                tmp$ratio_cat == "higher_in_manip"],1,-1)
## data for lower_in_manip
tmp_data_l = tibble("OTU"= taxa_names(ps_16S_fltr))
tmp_data_l[,"susp_C"] <- -1
tmp_data_l[,"susp_MAC_a1"] <- ifelse(tmp_data_l$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_MAC_a1" &
                                                tmp$manip_susp_relab >0 &
                                                tmp$ratio_cat == "lower_in_manip"],1,-1)
tmp_data_l[,"susp_MAC_a2"] <- ifelse(tmp_data_l$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_MAC_a2" &
                                                tmp$manip_susp_relab >0 &
                                                tmp$ratio_cat == "lower_in_manip"],1,-1)
tmp_data_l[,"susp_PEB_a1"] <- ifelse(tmp_data_l$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_PEB_a1" &
                                                tmp$manip_susp_relab >0 &
                                                tmp$ratio_cat == "lower_in_manip"],1,-1)
tmp_data_l[,"susp_PEB_a2"] <- ifelse(tmp_data_l$OTU %in% 
                                     tmp$OTU[tmp$manip_susp == "susp_PEB_a2" &
                                                tmp$manip_susp_relab >0 &
                                                tmp$ratio_cat == "lower_in_manip"],1,-1)
# extract data
write.csv(tmp_data,"iTOL/tmp_itol_ns.csv",row.names = F,quote = F)
write.csv(tmp_data_h,"iTOL/tmp_itol_h.csv",row.names = F,quote = F)
write.csv(tmp_data_l,"iTOL/tmp_itol_l.csv",row.names = F,quote = F)


```


## effect of coal (stat_q2)
Effect of coalescence on the 236 most abundant OTUs detected in suspensions:

```{r plot by rel ab in susp}

# combine with stat_q2 data
tmp_stat_q2 <- stat_q2_otu_comparison_sum_w0 %>%
  dplyr::filter(OTU %in% stat_susp_OTU$OTU 
                & pval_adjust <= 0.05) %>%
  select(OTU,g2,estimate) %>%
  pivot_wider(.,names_from = g2, values_from = estimate) %>%
  left_join(tmp_relab_susp,.)
tmp_stat_q2[is.na(tmp_stat_q2)] <- 0

# wrangle data
tmp_name_d = c("d1","d2","d3")
tmp_loop_df = tibble()

for (tmp_i in tmp_name_susp) {
  for (tmp_j in tmp_name_d) {
    
    tmp_df <- tibble("OTU"=tmp_stat_q2$OTU,
                     "susp_C"=unlist(tmp_stat_q2$susp_C),
                     "susp"=tmp_i,
                     "manip_susp_relab"=unlist(tmp_stat_q2[,paste0("susp_",tmp_i)]),
                     "ratio"= unlist(tmp_stat_q2[,tmp_i]),
                     "treatment"=paste0(tmp_i,tmp_j),
                     "estimate"= unlist(tmp_stat_q2[,paste0(tmp_i,tmp_j)]))
    tmp_loop_df <- rbind(tmp_loop_df,tmp_df)

    
  }
  
}

## in the way the ratio was calculated, a ratio >0.5 means higher rel ab inn manip susp
tmp_loop_df[,"ratio_cat"] <- ifelse(tmp_loop_df$ratio >0.5,"higher_in_manip","lower_in_manip")


# PLOT REL AB IN SUSP
for (tmp_i in tmp_name_susp) {
  print(
    tmp_loop_df %>%
      filter(susp %in% tmp_i) %>%
      ggplot() +
      aes(x = susp_C, y = manip_susp_relab, fill = -estimate, size = abs(estimate)) +
      geom_point(shape = "circle filled") +
      geom_abline(intercept = 0) +
      facet_wrap(vars(treatment)) +
      scale_x_continuous(trans = "asn") +
      scale_y_continuous(trans = "asn") +
      scale_fill_gradient2(low = "#3399FF" ,mid = "white",high = "#FF3366",
                           limits = c(-6, 6)) +
      scale_size_continuous(limits = c(0, 6)) +
      theme_light() +
      labs(title = tmp_i)
    )
}

# TEST CORRELATION

## should I test the correlation only for OTUs detected in the manipulated suspensions?
tmp_cor_df <- tmp_loop_df #%>%
  #filter(OTU %in% tmp_OTUs_56)

# use the lm function in R to fit a linear regression model and extract the formula
# I calculate correlation for each treatment to not replicate the suspension ratio data
for (tmp_i in tmp_name_susp) {
  for (tmp_j in tmp_name_d) {
    tmp_df <- tmp_cor_df %>%
      filter(susp %in% tmp_i
             & treatment %in% paste0(tmp_i,tmp_j))
    tmp_aov <- aov(estimate ~ ratio_cat, data = tmp_df)
    tmp_aov_sum <- summary(tmp_aov)
    
    print(paste0(tmp_i,tmp_j))
    print(format(tmp_aov_sum[[1]][["Pr(>F)"]][1]*12, digits = 3,scientific = T))
    print(ifelse(tmp_aov_sum[[1]][["Pr(>F)"]][1]*12 > 0.05,"n.s.",
                 ifelse(tmp_aov_sum[[1]][["Pr(>F)"]][1]*12 > 0.01,"*",
                        ifelse(tmp_aov_sum[[1]][["Pr(>F)"]][1]*12 > 0.001,"**",
                               "***"
                               ))))

  }
}

# test: estimate ~ susp*ratio_cat
tmp_aov <- aov(estimate ~ susp*ratio_cat*treatment, data = tmp_cor_df)
tmp_aov_sum <- summary(tmp_aov)
tmp_aov_sum
length(unique(tmp_cor_df$OTU))

# boxplot of the two variables
for (tmp_i in tmp_name_susp) {
  print(
    tmp_cor_df %>%
      filter(susp %in% tmp_i) %>%
      ggplot() +
      aes(x = ratio_cat, y = -estimate) +
      geom_boxplot() +
      geom_point(shape = "circle filled", aes(fill = -estimate, size = abs(estimate))) +
      facet_wrap(vars(treatment)) +
      scale_y_continuous(limits = c(-6, 6)) +
      scale_fill_gradient2(low = "#3399FF" ,mid = "white",high = "#FF3366",
                           limits = c(-6, 6)) +
      scale_size_continuous(limits = c(0, 6)) +
      theme_light()
    )
  
}





```

```{r plot by coalescence effect}

# stat q2
tmp_stat_q2 <- stat_q2_otu_comparison_sum_w0 %>%
  dplyr::filter(OTU %in% stat_susp_OTU$OTU 
                & pval_adjust <= 0.05) %>%
  select(OTU,g2,estimate)
# stat susp
tmp_data <- stat_susp_OTU %>%
  select(OTU,manip_susp,ratio_cat)

# combine with stat_q2 data
tmp_loop_df <- tibble()
tmp_name_susp = c("MAC_a1","MAC_a2","PEB_a1","PEB_a2")
for (tmp_i in unique(stat_susp_OTU$OTU)) {
  for (tmp_j in tmp_name_susp) {
    tmp = tibble("OTU"=tmp_i,
                 "manip_susp"=paste0("susp_",tmp_j),
                 "treatment"=paste0(tmp_j,c("d1","d2","d3")))
    tmp <- left_join(tmp,tmp_data,by=c("OTU","manip_susp"))
    tmp <- left_join(tmp,tmp_stat_q2,by=c("OTU","treatment"="g2"))
    tmp_loop_df <- rbind(tmp_loop_df,tmp)
  }
}
tmp_loop_df$estimate[is.na(tmp_loop_df$estimate)] <- 0
tmp_loop_df[,"sign_comp"] <- ifelse(tmp_loop_df$estimate < 0,"up",
                                    ifelse(tmp_loop_df$estimate > 0,"down","n.s."))
tmp_loop_df[,"ratio_cat2"] <- ifelse(tmp_loop_df$ratio_cat == "higher_in_manip",
                                     "higher_in_manip","other")

# plot
## factor ordering and colors
tmp_loop_df$treatment <- factor(tmp_loop_df$treatment,
                     levels = rev(sort(unique(tmp_loop_df$treatment))))
tmp_loop_df$sign_comp <- factor(tmp_loop_df$sign_comp,levels = c("up","down","n.s."))

## boxplot
ggplot(tmp_loop_df) +
  aes(x = -estimate, y = treatment, fill = ratio_cat) +
  geom_boxplot() +
  geom_point(shape="circle filled",position = position_dodge(width = 0.75))+
  scale_fill_manual(values = c("#FF3366","#3399FF","#333333"))  +
  theme_light()

## barplot
tmp_loop_df %>%
  dplyr::filter(!(sign_comp %in% "n.s.")) %>%
  ggplot() +
  aes(y = treatment, fill = ratio_cat) +
  geom_bar() +
  scale_fill_manual(values = c("#FF3366","#3399FF","#333333")) +
  theme_light() +
  facet_wrap(vars(sign_comp))

# test: estimate ~ susp*ratio_cat
tmp_aov <- aov(estimate ~ manip_susp*ratio_cat*treatment, data = tmp_loop_df)
tmp_aov_sum <- summary(tmp_aov)
tmp_aov_sum
## post hoc Tukey test
library(agricolae)
tmp_comp <- HSD.test(tmp_aov,c('ratio_cat','treatment'),alpha = 0.05,group = T)
## tibble with statistical groups
tmp_stat = tibble("ratio_cat"=rownames(tmp_comp[["groups"]]),
                  "mean"=tmp_comp[["groups"]][["tmp_data$value"]],
                  "stat_groups"=as.character(tmp_comp[["groups"]][["groups"]]))



```



## effect of properties (stat_q1)

```{r effect of properties (236 OTUs)}

# /!\ need to run {r OTUs in susp}

tmp_stat_q1 <- stat_q1_otu_rel %>%
  dplyr::filter(OTU %in% stat_susp_OTU$OTU &
                  var != "Residuals")
tmp_stat_q1[,"ratio_cat"] <- ifelse(tmp_stat_q1$OTU %in%
                                      unique(stat_susp_OTU$OTU[stat_susp_OTU$ratio_cat == "higher_in_manip"]),
                                    "higher_in_manip",
                                    ifelse(tmp_stat_q1$OTU %in% 
                                             unique(stat_susp_OTU$OTU[stat_susp_OTU$ratio_cat == "lower_in_manip"]),
                                    "lower_in_manip","n.s."))

# test: F-value ~ var*ratio_cat
tmp_aov <- aov(rel_fval ~ var*ratio_cat, data = tmp_stat_q1)
tmp_aov_sum <- summary(tmp_aov)
tmp_aov_sum

# plot
## factor ordering and colors
tmp_stat_q1$var <- factor(tmp_stat_q1$var,
                     levels = rev(c("a","b","d","b:a","a:d","b:d","b:a:d")))
tmp_stat_q1$var_type <- factor(tmp_stat_q1$var_type,
                     levels = c("main","double","triple"))
tmp_colors <- rev(c("#9b2226","#db912b","#005f73","#bb3e03","#ecba53","#0a9396","#94d2bd"))
## plot relative property effects
tmp_stat_q1 %>%
  filter(pval_adjust <= 0.05) %>% 
  dplyr::group_by(ratio_cat,var,var_type) %>%
  dplyr::summarise(mean_rfval = mean(rel_fval)) %>%
  ggplot() +
  aes(x= ratio_cat, y= mean_rfval, fill= var)+
  geom_col(position = "fill")+
  scale_fill_manual(values = tmp_colors)+
  coord_flip()+
  facet_grid(ratio_cat~var_type,
             scales = "free",
             switch = "y", 
             space = "free_y") +
  theme_classic() +
    theme(strip.text = element_text(face="bold", size=10),
        strip.background = element_rect(fill="#ffffff", colour="#ffffff",size=1.5),
        strip.placement = "outside")


```

## q1 & q2


```{r by treatment}

# which OTUs are affected by coal
tmp_df1 = stat_q2_otu_comparison_sum_w0 %>% 
  filter(pval_adjust <= 0.05 
         & OTU %in% tmp_OTUs_56)
tmp_df1[,"sign_comp"] <- ifelse(tmp_df1$estimate <0,"1up","2down") # /!\ up = estimates <0
tmp_df1 <- tmp_df1[,colnames(tmp_df1)[c(1,9,12)]]

# which OTUs have signif fval
tmp_df2 <- stat_q1_otu %>% 
  filter(pval_adjust <= 0.05
         & OTU %in% tmp_OTUs_56)
tmp_df2 <- tmp_df2[,colnames(tmp_df2)[c(1,2,6)]]

# how many sign_comp by var
tmp_df <- left_join(tmp_df1,tmp_df2,by = "OTU")
tmp_df <- na.omit(tmp_df)

tmp_df <- tmp_df %>%
  dplyr::group_by(g2,var,sign_comp) %>%
  dplyr::summarise(nb_comp = n())
tmp_df

# add nb OTUs by sign_comp by ttt
tmp_df3 <- tmp_df1 %>%
  dplyr::group_by(g2,sign_comp) %>%
  dplyr::summarise(nb_OTUs_heatmap = n()) # divide by nb var to have total nb of OTUs
#  dplyr::summarise(nb_OTUs_heatmap = n()/length(unique(tmp_df$var))) # divide by nb var to have total nb of OTUs
#tmp_df <- left_join(tmp_df,tmp_df3,by = c("g2", "sign_comp"))

# add nb OTUs by sign_comp by ttt affected by factor
tmp_df4 <- tmp_df1 %>% 
  filter(OTU %in% unique(tmp_df2$OTU)) %>%
  dplyr::group_by(g2,sign_comp) %>%
  dplyr::summarise(nb_OTUs_fval = n()) # divided by nb var to have total nb of OTUs
#  dplyr::summarise(nb_OTUs_fval = n()/length(unique(tmp_df$var))) # divided by nb var to have total nb of OTUs
#tmp_df <- left_join(tmp_df,tmp_df4,by = c("g2", "sign_comp"))

# plot
## factor ordering and colors
tmp_df3$g2 <- factor(tmp_df3$g2,
                     levels = rev(sort(unique(tmp_df3$g2))))
tmp_df$var <- factor(tmp_df$var,
                     levels = rev(c("a","b","d","b:a","a:d","b:d","b:a:d")))
tmp_colors <- rev(c("#9b2226","#db912b","#005f73","#bb3e03","#ecba53","#0a9396","#94d2bd"))


ggplot() +
  geom_col(data = tmp_df3, aes(x = g2,y=nb_OTUs_heatmap),fill="#cccccc") +
  geom_col(data = tmp_df4, aes(x = g2,y=nb_OTUs_fval),fill="#999999") +
  geom_col(data = tmp_df,aes(x = g2,y=nb_comp, fill = var),position = position_dodge(preserve = "single")) +
  scale_fill_manual(values = tmp_colors) +
  coord_flip() +
  theme_light() +
  facet_wrap(vars(sign_comp))

# check
tmp_df[tmp_df$g2 == "MAC_a1d1" & tmp_df$var == "d",]
tmp_OTUs1 = tmp_df1$OTU[tmp_df1$g2 == "MAC_a1d1" & tmp_df1$sign_comp == "down"]
tmp_OTUs2 = tmp_df2$OTU[tmp_df2$var == "d"]
tmp_OTUs3 = unique(tmp_df2$OTU)
length(intersect(tmp_OTUs1,tmp_OTUs3))

```

## network

```{r OTUs in network}

tmp_edges_otu <- PLN_model.StARS_xp2_edges
tmp_nodes = PLN_model.StARS_xp2_nodes

tmp_OTUs_susp = unique(stat_susp_OTU$OTU)
tmp_OTUs_susp_higher = unique(stat_susp_OTU$OTU[stat_susp_OTU$ratio_cat == "higher_in_manip"])

# are OTUs in network?
table(tmp_OTUs_susp %in% tmp_nodes)
table(tmp_OTUs_susp_higher %in% tmp_nodes)

# how many edges?
table(tmp_edges_otu$source %in% tmp_OTUs_susp | tmp_edges_otu$target %in% tmp_OTUs_susp)
table(tmp_edges_otu$source %in% tmp_OTUs_susp_higher | tmp_edges_otu$target %in% tmp_OTUs_susp_higher)


```

```{r edges with tmp_OTUs_susp_higher}

tmp_OTUs_susp_higher = unique(stat_susp_OTU$OTU[stat_susp_OTU$ratio_cat == "higher_in_manip"])

tmp_edges_otu <- PLN_model.StARS_xp2_edges %>%
  dplyr::filter(source %in% tmp_OTUs_susp_higher | target %in% tmp_OTUs_susp_higher)

# how many pos as FALSE / neg as TRUE
table(tmp_edges_otu$value <0)

# how many links by OTU?
tmp_top_OTUs = sort(table(c(tmp_edges_otu$source,tmp_edges_otu$target)),decreasing = T)
tmp_top_OTUs <- tmp_top_OTUs[names(tmp_top_OTUs) %in% tmp_OTUs_susp_higher]
tmp_top_OTUs

# most connected OTUs
tmp_edges_otu <- PLN_model.StARS_xp2_edges %>%
  filter(source %in% names(tmp_top_OTUs)[1:2] | target %in% names(tmp_top_OTUs)[1:2])
## how many pos as FALSE / neg as TRUE for 1 OTU
table(tmp_edges_otu$value[tmp_edges_otu$source %in% names(tmp_top_OTUs)[2] 
                          | tmp_edges_otu$target %in% names(tmp_top_OTUs)[2]] <0)


```

### venn


```{r venn & top OTUs}



# /!\ need tmp_venn_edges_taxo from {r taxo_by_edges} in 6.1.Compare_network_stat.Rmd
tmp_OTUs = c("OTU-219","OTU-45")

tmp_venn_top <- tmp_venn_edges_taxo %>%
  filter(source %in% tmp_OTUs | target %in% tmp_OTUs)
tmp_venn_top[,"top_OTU"] <- ifelse(tmp_venn_top$source == "OTU-219","OTU-219",
                                   ifelse(tmp_venn_top$target == "OTU-219","OTU-219",
                                          "OTU-45"))

# plot nb link by factor
## reorder factor + colors
tmp_venn_top$set <- factor(tmp_venn_top$set,
                             levels = rev(c("all..compo..dens","all..div..dens","all..div..compo",
                                        "all..dens","all..compo","all..div","all","all..div..compo..dens")))
tmp_colors <- rev(c("#9b2226","#db912b","#005f73","#bb3e03","#ecba53","#0a9396","#94d2bd","#666666"))
## plot
tmp <- tmp_venn_top %>%
  dplyr::group_by(set,edge_type,top_OTU) %>%
  dplyr::summarise(count=n())
tmp$count <- ifelse(tmp$edge_type =="pos",tmp$count,-tmp$count)
tmp %>%
  ggplot() +
  aes(x = set, y = count, fill = set) +
  geom_col() +
  geom_hline(yintercept = 0,color = "#333333") +
  ylim(c(-10,10)) +
  scale_fill_manual(values = tmp_colors) +
  coord_flip() +
  facet_wrap(vars(top_OTU)) +
  theme_light()

# plot nb links with factor by taxa
tmp <- tmp_venn_top %>%
  dplyr::group_by(taxa,set,edge_type,top_OTU) %>%
  dplyr::summarise(count=n())
tmp$count <- ifelse(tmp$edge_type =="pos",tmp$count,-tmp$count)
tmp %>%
  ggplot() +
  aes(x = taxa, y = count, fill = set) +
  geom_col() +
  geom_hline(yintercept = 0,color = "#333333") +
  ylim(c(-7,7)) +
  scale_fill_manual(values = tmp_colors) +
  coord_flip() +
  facet_grid(top_OTU~.,scales = "free") +
  theme_light() 


```

```{r plot several OTU abundance}

# /!\ need tmp_venn_top from {r venn & top OTUs}

tmp_name_edge = "Bacillus - Pseudochrobactrum"
#tmp_name_edge = "Massilia - Stenotrophomonas"

tmp_OTUs <- sort(unique(c(tmp_venn_top$source[tmp_venn_top$taxa == tmp_name_edge],
                          tmp_venn_top$target[tmp_venn_top$taxa == tmp_name_edge])))

tmp_ps = ps_16S_fltr
## transform to relative abundance 
tmp_ps <- transform_sample_counts(tmp_ps,function(x) x / sum(x))
## prune taxa
tmp_ps = prune_taxa(taxa_names(tmp_ps) %in% tmp_OTUs,tmp_ps)

tmp_data = psmelt(tmp_ps)
# treatment colors
tmp_colors <- Color_sets[Color_sets$set == "treatments",c("treatment","color","order")]  
tmp_colors <- tmp_colors[tmp_colors$treatment %in% tmp_data$treatment,]

ggplot(tmp_data,aes(tmp_data$treatment,tmp_data$Abundance,
                         fill=treatment))+
  geom_boxplot(outlier.size = 0.1,lwd=0.1)+
  geom_point(color="#333333", size=0.1) +
  labs(title = "Node OTUs",y="Relative abundance (among filtered OTUs)",x="Treatments")+
#  labs(title = tmp_OTUs,y="Raw abundance (log10 scale)",x="Treatments")+
  #scale_y_log10()+
  scale_y_continuous(trans = "asn") +
  scale_fill_manual(values = tmp_colors$color)+
  #geom_hline(yintercept = 0.25/100)+
  #geom_hline(yintercept = 0.1/100)+
  geom_hline(yintercept = 0)+
  facet_grid(.~Genus+OTU)+
  theme_light()+
  theme(axis.text.x  = element_text(vjust=0.5, size=12,angle = 90))

view(tmp_abund[tmp_abund$OTU %in% tmp_OTUs,])

```

## rank curve

```{r rank curve}

# /!\ need tmp_relab_susp in {r OTUs in susp}

tmp_nodes = PLN_model.StARS_xp2_nodes

for (tmp_i in tmp_name_susp) {
  tmp_abund <- tibble("OTU"=tmp_relab_susp$OTU,
                      "susp_C"=tmp_relab_susp$susp_C,
                      "manip_susp"=unlist(tmp_relab_susp[,paste0("susp_",tmp_i)]))
  # set in or out network
  tmp_abund[,"network"] <- ifelse(tmp_abund$OTU %in% tmp_nodes,"In network","Not in network")
  
  # plot
  print(
    ggplot(tmp_abund)+
      geom_col(aes(x=reorder(OTU,desc(susp_C)),y=susp_C),fill = "#cccccc")+
      geom_col(aes(x=reorder(OTU,desc(susp_C)),y=manip_susp,fill = network))+
      geom_vline(xintercept = seq(10.5,ntaxa(tmp_ps),10),lwd=0.1, color = "#cccccc") +
      scale_y_continuous(trans = "asn") +
      scale_fill_manual(name = "Presence in network",values = c("#FF6666","#333333"))+
      labs(title = tmp_i) +
      theme_light() +
      theme(axis.text.x  = element_text(vjust=0.5, size=5,angle = 90),
        panel.grid.major.x = element_blank())
    )
  
}


colnames(tmp_abund)

# plot
ggplot(tmp_abund)+
  geom_col(aes(x=reorder(OTU,desc(susp_C)),
                                 y=susp_PEB_a2,
                                 fill=network))+
  scale_y_continuous(trans = "asn") +
  theme_light() +
  theme(axis.text.x  = element_text(vjust=0.5, size=5,angle = 90),
        panel.grid.major.x = element_line(color = "white", size = 0.8))+
  scale_fill_manual(values = c("#FF6666","#666666"))+
  geom_vline(xintercept = seq(10.5,ntaxa(tmp_ps),10),
             lwd=0.1, color = "#cccccc")

#```

#```{r extract fig}
library(ggedit)

p[["labels"]][["title"]] <- element_blank()
p[["labels"]][["x"]] <- element_blank()
p[["labels"]][["y"]] <- element_blank()
p[["theme"]][["legend.position"]] <- 'none'
p[["theme"]][["axis.text.x"]] <- element_blank()
p[["theme"]][["axis.text.y"]] <- element_blank()
# export svg
ggsave(plot = p,
       filename = "C:/Users/srhuet/Downloads/tmp_fig.svg",
       width = 156,height = 36,units = "mm")

```

### bin ?

```{r OTUs in susp}

# calculate rel_ab in susp
tmp_ps = ps_16S_susp
## remove T0 samples
tmp_ps <- prune_samples(tmp_ps@sam_data$treatment != "T0",tmp_ps)
tmp_ps <- prune_taxa(taxa_sums(tmp_ps) >0,tmp_ps)

## keep OTUs among most abund in microcosms
tmp_ps <- prune_taxa(taxa_names(tmp_ps) %in% taxa_names(ps_16S_fltr),tmp_ps)

## transform to relative abundance 
tmp_ps <- transform_sample_counts(tmp_ps,function(x) x / sum(x))
## extract data
tmp_relab_susp = psmelt(tmp_ps) %>%
  
  ## keep OTUs among most abund in microcosms: did before calculating the rel ab
  #dplyr::filter(OTU %in% taxa_names(ps_16S_fltr)) %>% 
  
  ## calculate median rel abund for each OTUs in each susp
  dplyr::group_by(OTU,treatment) %>%
  dplyr::summarise(mean_relab = mean(Abundance)) %>%
  pivot_wider(.,names_from = treatment, values_from = mean_relab) %>%
  ungroup()

# calculate rel abund ratio as: manip susp / (manip susp + susp_C)
tmp_name_susp = c("MAC_a1","MAC_a2","PEB_a1","PEB_a2")
for (tmp_i in tmp_name_susp) {
  tmp_relab_susp[,tmp_i] <- tmp_relab_susp[,paste0("susp_",tmp_i)] /
    (tmp_relab_susp[,paste0("susp_",tmp_i)] + tmp_relab_susp$susp_C)
}

# pivot longer
tmp_relab_ratio = tibble()
for (tmp_i in tmp_name_susp) {
  tmp_df <- tibble("OTU"=tmp_relab_susp$OTU,
                   "susp_C"=unlist(tmp_relab_susp$susp_C),
                   "susp"=tmp_i,
                   "manip_susp_relab"=unlist(tmp_relab_susp[,paste0("susp_",tmp_i)]),
                   "ratio"= unlist(tmp_relab_susp[,tmp_i]))
  tmp_relab_ratio <- rbind(tmp_relab_ratio,tmp_df)

}
# add cat var for ratio
tmp_relab_ratio[is.na(tmp_relab_ratio)] <- 0
## in the way the ratio was calculated, a ratio >0.5 means higher rel ab inn manip susp
tmp_relab_ratio[,"ratio_cat"] <- ifelse(tmp_relab_ratio$ratio >0.5,"higher_in_manip","lower_in_manip")


# OTUs in manipulated susp and most abund in microcosms
tmp_OTUs_56 = intersect(taxa_names(ps_16S_susp_manip),taxa_names(ps_16S_fltr))
# which OTU higher in manip susp
tmp_OTUs_susp_higher <- unique(tmp_relab_ratio$OTU[tmp_relab_ratio$ratio >0.5])

```

```{r diff abund between susp}
library(lme4)
library(emmeans)

tmp_ps = ps_16S_susp
## remove T0 samples
tmp_ps <- prune_samples(tmp_ps@sam_data$treatment != "T0",tmp_ps)
tmp_ps <- prune_taxa(taxa_sums(tmp_ps) >0,tmp_ps)
## subset taxa
tmp_ps = prune_taxa(taxa_names(tmp_ps) %in% taxa_names(ps_16S_fltr),tmp_ps)


# define model variables
#  treatment
tmp_ttt = unlist(tmp_ps@sam_data$treatment)
names(tmp_ttt) <- unlist(tmp_ps@sam_data$sample)
# offset
tmp_o = log(sample_sums(tmp_ps))
names(tmp_o) <- unlist(tmp_ps@sam_data$sample)
# random effect
tmp_z <- unlist(tmp_ps@sam_data$sample)
  
# global df
tmp_model_sum_global = tibble()
tmp_emmeans_sum_global = tibble()


for (tmp_i in taxa_names(tmp_ps)) {
  tmp_OTU = tmp_i
  
  # response variable
  tmp_y = as.vector(tmp_ps@otu_table[tmp_OTU,]@.Data)
  
  tmp_glmer1 <- glmer(tmp_y ~ -1 + tmp_ttt + (1 | tmp_z),family='poisson', offset = tmp_o)
  tmp_glmer0 <- glmer(tmp_y ~ 1 + (1 | tmp_z),family='poisson', offset = tmp_o)
  
  # LRT
  tmp_LRT <- 2*(logLik(tmp_glmer1)[1] - logLik(tmp_glmer0)[1])
  ## calcultate Chisq pval and Bonferroni correction for 258 OTUs
  tmp_LRT_pchisq = pchisq(tmp_LRT, df=1, lower.tail=FALSE) *ntaxa(tmp_ps)

  if (tmp_LRT_pchisq <= 0.05) {
    
    # model summary
    tmp_model_sum = summary(tmp_glmer1)$coefficients
    tmp_model_sum = tibble("OTU"= tmp_OTU,
                           "treatment"=rownames(tmp_model_sum),
                           as_tibble(tmp_model_sum))
    
    # multiple comparaison
    tmp_emmeans = emmeans(tmp_glmer1,pairwise~tmp_ttt,adjust="none")
    ## select p value
    tmp_emmeans_sum = summary(tmp_emmeans)
    tmp_emmeans_sum = tmp_emmeans_sum[["contrasts"]]
    ## wrangle summary
    tmp_df = tmp_emmeans_sum
    tmp = unlist(strsplit(as.character(tmp_df$contrast)," - "))
    tmp_df[,"g1"] <- tmp[seq(1,length(tmp),by=2)]
    tmp_df[,"g2"] <- tmp[seq(2,length(tmp),by=2)]
    # tmp_df = tmp_df[tmp_df$g1 == "C",] # keep only comparison to control
    tmp_emmeans_sum = cbind("OTU"=tmp_OTU,tmp_df)
    
    rm(tmp_emmeans,tmp,tmp_df)
    
    } else {
      # empty model summary
      tmp_model_sum = tibble("OTU"=tmp_OTU,
                             "treatment"=NA,
                             "Estimate"=NA,
                             "Std. Error"=NA,
                             "z value"=NA,
                             "Pr(>|z|)"=1)

      
      # empty emmeans summary
      tmp_emmeans_sum = tibble("OTU"=tmp_OTU,
                               "contrast"=NA,
                               "estimate"=NA,
                               "SE"=NA,
                               "df"=NA,
                               "z.ratio"=NA,
                               "p.value"=1,
                               "g1"=NA,
                               "g2"=NA)
  }
  
  # global df
  tmp_model_sum_global = rbind(tmp_model_sum_global,tmp_model_sum)
  tmp_emmeans_sum_global = rbind(tmp_emmeans_sum_global,tmp_emmeans_sum)


}


# bonferroni p-val correction : 214 OTUs * 10 comparisons
tmp_emmeans_sum_global[,"pval_adjust"] <- tmp_emmeans_sum_global$p.value * ntaxa(tmp_ps) *10

# keep signif estimates
tmp_emmeans_sum_global[,"rel_estimate"] <- ifelse(tmp_emmeans_sum_global$pval_adjust <=0.05,
                                                  tmp_emmeans_sum_global$estimate,
                                                  0)

```

```{r edges}

tmp_edges_otu <- PLN_model.StARS_xp2_edges
# are connected OTUs detected in manip susp
tmp_edges_otu[,"source_in_manip_susp"] <- ifelse(tmp_edges_otu$source %in% tmp_OTUs_56,"yes","no")
tmp_edges_otu[,"target_in_manip_susp"] <- ifelse(tmp_edges_otu$target %in% tmp_OTUs_56,"yes","no")

# which edges
table(tmp_edges_otu$source %in% tmp_OTUs_56 & tmp_edges_otu$target %in% tmp_OTUs_56)
table(tmp_edges_otu$source %in% tmp_OTUs_20 | tmp_edges_otu$target %in% tmp_OTUs_20)

# most connected OTUs
tmp <- table(c(tmp_edges_otu$source[tmp_edges_otu$source_in_manip_susp =="yes"],
        tmp_edges_otu$target[tmp_edges_otu$target_in_manip_susp =="yes"]))
tmp <- tmp[order(tmp,decreasing = T)][1:4]

tmp_most_connect_otu <- tmp_edges_otu %>%
  dplyr::filter(source %in% names(tmp)
                | target %in% names(tmp))


# plot with genus
tmp_plot_data <- tmp_df_most_lkd %>%
  dplyr::filter(edge_name %in% tmp_most_connect_otu$edge_name)

tmp_colors <- c("#9b2226","#db912b","#005f73","#bb3e03","#ecba53","#0a9396","#94d2bd","#666666")

## plot nb links with factor by taxa
tmp <- tmp_plot_data %>%
  dplyr::filter(source_in_manip_susp == "yes" | target_in_manip_susp == "yes") %>%
  dplyr::group_by(taxa,set,edge_type) %>%
  dplyr::summarise(count=n())
tmp$count <- ifelse(tmp$edge_type =="pos",tmp$count,-tmp$count)
tmp %>%
  ggplot() +
  aes(x = taxa, y = count, fill = set) +
  geom_col() +
  scale_fill_manual(values = tmp_colors) +
  theme_light() +
  coord_flip() +
  geom_hline(yintercept = 0,color = "#333333") +
  ylim(c(-7,7))

```


```{r most linked genus}

# /!\ see {r plot most linked genus} in 6.1.Compare_network_stat.Rmd

# are most connected genus detected in manip susp
tmp_df_most_lkd[,"source_in_manip_susp"] <- ifelse(tmp_df_most_lkd$source %in% tmp_OTUs_56,"yes","no")
tmp_df_most_lkd[,"target_in_manip_susp"] <- ifelse(tmp_df_most_lkd$target %in% tmp_OTUs_56,"yes","no")

# which edges
table(tmp_df_most_lkd$source_in_manip_susp == "yes" | tmp_df_most_lkd$target_in_manip_susp == "yes")
# which OTUs
intersect(unique(c(tmp_df_most_lkd$source,tmp_df_most_lkd$target)),tmp_OTUs_56)
# from which genus?
taxtab$genus_clean[taxtab$OTU %in% intersect(unique(c(tmp_df_most_lkd$source,tmp_df_most_lkd$target)),tmp_OTUs_56)]


# plot
tmp_colors <- c("#9b2226","#db912b","#005f73","#bb3e03","#ecba53","#0a9396","#94d2bd","#666666")

## plot nb links with factor by taxa
tmp <- tmp_df_most_lkd %>%
  dplyr::filter(source_in_manip_susp == "yes" | target_in_manip_susp == "yes") %>%
  dplyr::group_by(taxa,set,edge_type) %>%
  dplyr::summarise(count=n())
tmp$count <- ifelse(tmp$edge_type =="pos",tmp$count,-tmp$count)
tmp %>%
  ggplot() +
  aes(x = taxa, y = count, fill = set) +
  geom_col() +
  scale_fill_manual(values = tmp_colors) +
  theme_light() +
  coord_flip() +
  geom_hline(yintercept = 0,color = "#333333") +
  ylim(c(-7,7))


```




























